### 2023-02-26 main.tex
customize template

\documentclass{INTERSPEECH2023}

% **************************************
% *    DOUBLE-BLIND REVIEW SETTINGS    *
% **************************************
% Comment out \interspeechcameraready when submitting the 
% paper for review.
% If your paper is accepted, uncomment this to produce the
%  'camera ready' version to submit for publication.
\interspeechcameraready

% imports

\usepackage{siunitx}

\title{Comparative Analysis of the wav2vec 2.0 Feature Extractor}
\name{Peter Vieting$^{1}$, Ralf Schl\"uter$^{1,2}$, Hermann Ney$^{1,2}$}
\address{
  $^1$RWTH Aachen University, Germany\\
  $^2$AppTek GmbH, Germany}
\email{\{vieting,schlueter,ney\}@cs.rwth-aachen.de}

\begin{document}

\maketitle

\begin{abstract}
\end{abstract}
\noindent\textbf{Index Terms}: speech recognition, feature extraction, raw waveform modeling

\section{Introduction}

\section{Related work}

\section{Methods}

\section{Results}

\section{Conclusions}

\section{Acknowledgements}

\ifinterspeechfinal
     Personal information revealing author identity
\else
     Anonymized information not revealing author identity
\fi

\bibliographystyle{IEEEtran}
\bibliography{mybib}

\end{document}

### 2023-02-26 acronyms.tex
add my acronyms and commands

% Acronym definitions.
% --------------------
% To disable certain acronyms without removing them everywhere, see this instructions:
% https://tex.stackexchange.com/q/424462/7396
%A
\newacronym{AM}{AM}{acoustic model}
\newacronym{AMI}{AMI}{Augmented Multi-party Interaction}
\newacronym{ARQ}{ARQ}{automatic repeat request}
\newacronym{ASR}{ASR}{automatic speech recognition}
%B
\newacronym[longplural={bi-directional long-short term memories}]{BLSTM}{BLSTM}{bi-directional long-short term memory}
\newacronym{BSS}{BSS}{blind speech separation}
%C
\newacronym{CART}{CART}{classification and regression tree}
\newacronym{CE}{CE}{cross entropy}
\newacronym{CER}{CER}{character error rate}
\newacronym{CDp}{CDp}{context dependent phoneme}
\newacronym{CNN}{CNN}{convolutional neural network}
\newacronym{CPC}{CPC}{contrastive predictive coding}
\newacronym{CTC}{CTC}{connectionist temporal classification}
%D
\newacronym{DCT}{DCT}{discrete cosine transform}
\newacronym{DL}{DL}{deep learning}
\newacronym{DNN}{DNN}{deep neural network}
\newacronym{DNN-HMM}{DNN-HMM}{deep neural network hidden Markov model}
%E
\newacronym{ELBO}{ELBO}{evidence lower bound}
\newacronym{EM}{EM}{expectation maximization}
%F
\newacronym{FE}{FE}{feature extractor}
\newacronym{FIR}{FIR}{finite impulse response}
\newacronym{FFNN}{FFNN}{feed-forward neural network}
\newacronym{fCE}{fCE}{frame-wise cross-entropy}
%G
\newacronym{G2P}{G2P}{grapheme-to-phoneme conversion}
\newacronym{GAN}{GAN}{generative adversarial network}
\newacronym{GMM}{GMM}{Gaussian mixture model}
\newacronym{GMM-HMM}{GMM-HMM}{Gaussian mixture model hidden Markov model}
\newacronym{GPU}{GPU}{graphics processing unit}
%H
\newacronym{HMM}{HMM}{hidden Markov model}
%I
\newacronym{IHM}{IHM}{individual headset microphones}
\newacronym{IIR}{IIR}{infinite impulse response}
%L
\newacronym{LAS}{LAS}{listen-attend-spell}
\newacronym{LC-BLSTM}{LC-BLSTM}{latency-controlled bidirectional long-short term memory}
\newacronym{LDA}{LDA}{linear discriminant analysis}
\newacronym{LM}{LM}{language model}
\newacronym[longplural={long-short term memories}]{LSTM}{LSTM}{long-short term memory}
%M
\newacronym{MDM}{MDM}{multiple distant microphones}
\newacronym{MFCC}{MFCC}{Mel-frequency cepstral coefficients}
\newacronym{MHSA}{MHSA}{multi-head self-attention}
\newacronym{MRES}{MRES}{multi-resolutional}
\newacronym{MT}{MT}{machine translation}
%N
\newacronym{NLP}{NLP}{natural language processing}
\newacronym{NN}{NN}{neural network}
%O
\newacronym{OOV}{OOV}{out-of-vocabulary}
%P
\newacronym{PC2}{$\text{PC}^{\text{2}}$}{Paderborn Center for Parallel Computing}

%R
\newacronym{RNA}{RNA}{recurrent neural aligner}
\newacronym{RNN}{RNN}{recurrent neural network}
\newacronym{RSAN}{RSAN}{recurrent selective attention network}
%S
\newacronym{SAT}{SAT}{speaker adaptive training}
\newacronym{SDM}{SDM}{single distant microphone}
\newacronym{SDR}{SDR}{signal-to-distortion ratio}
\newacronym{SC}{SC}{supervised convolutional}
\newacronym{sMBR}{sMBR}{state-level minimum Bayes risk}
\newacronym{STFT}{STFT}{short time Fourier transform}
%T
\newacronym{TDNN}{TDNN}{time delay neural network}
\newacronym[longplural={time-frequencies}]{tf}{tf}{time-frequency}
%U
%V
\newacronym{VAD}{VAD}{voice activity detection}
\newacronym{VTLN}{VTLN}{vocal tract length normalization}
%W
\newacronym{cpWER}{cpWER}{concatenated minimum-permutation word error rate}
\newacronym{WER}{WER}{word error rate}
\newacronym{WERR}{WERR}{word error rate reduction}
\newacronym{WP}{WP}{work package}
\newacronym{WPE}{WPE}{weighted prediction error}
\newacronym{WSJ}{WSJ}{Wall Street Journal}

### 2023-02-26 tables/features_pretraining.tex
add tables

\begin{table}[htbp]

\centering
\caption{Effect of pre-training the wav2vec 2.0 feature extractor using the unsupervised loss on the same data (Librispeech 960h)}
\label{table:features_pretraining}
\begin{tabular}{|c|c|S|}
\hline
\multicolumn{2}{|c|}{Feature Extractor} &       {WER} \\\hline
                              Trainable & Pretrained & {dev-other} \\\hline
                                    yes &         no &         6.8 \\\cline{2-3}
                                        &        yes &         6.9 \\\cline{1-1}\cline{3-3}
                                     no &            &         7.1 \\
\hline
\end{tabular}

\end{table}

### 2023-02-26 tables/create_tables.sh

@@ -1,0 +1,16 @@
+#!/usr/bin/env bash
+HOSTNAME=$(hostname)
+if [ "$HOSTNAME" == "ProBook" ]
+then
+    EXCEL2LATEX=/mnt/d/work/code/excel-to-latex/excel_to_latex/excel_to_latex.py
+else
+    EXCEL2LATEX=/u/vieting/Documents/text/excel-to-latex/excel_to_latex/excel_to_latex.py
+fi
+
+python ${EXCEL2LATEX} --file_in features.xlsx --all_lines --vertical_merge --sheet general --wer_cols E
+python ${EXCEL2LATEX} --file_in features.xlsx --all_lines --vertical_merge --sheet pretraining --wer_cols C
+python ${EXCEL2LATEX} --file_in features.xlsx --all_lines --vertical_merge --sheet window_size --wer_cols C
+python ${EXCEL2LATEX} --file_in features.xlsx --all_lines --vertical_merge --sheet i6_size --wer_cols E
+python ${EXCEL2LATEX} --file_in features.xlsx --all_lines --vertical_merge --sheet w2v_size --wer_cols D
+python ${EXCEL2LATEX} --file_in features.xlsx --all_lines --vertical_merge --sheet w2v_proj --wer_cols C
+

### 2023-02-26 main.tex
add my acronyms and commands
add content from draft

@@
 % imports
+\usepackage[acronym,toc,shortcuts,nonumberlist]{glossaries}
+\usepackage{multirow}
+\usepackage{siunitx}
+\usepackage{xcolor}
 
-\usepackage{siunitx}
+\input{acronyms}
+\input{commands}
 
@@
 \section{Methods}
+\subsection{Feature Extractors}
+In this work, we compare different feature extraction methods -- namely log Mel filterbank features, Gammatone features \cite{schlueter:icassp07}, \gls{SCF} features \cite{tuske2018:waveform} and the \gls{FE} of the \wvtwo model \cite{facebook2020wav2vec2}.
+
+\subsection{Model Sizes}
+By default, I use 40k parameters for i6 features compared to 4M for \wvtwo.
+Can we improve the performance by making the model larger?
+\subsubsection{wav2vec 2.0 Components}
+To understand where \wvtwo uses the parameters and how much they contribute to the \gls{FE}'s performance, we run it with different configurations.
+The results for different widths and depths are shown in \refTab{table:features_w2v_size}.
+% Kernel sizes and strides were chosen as (10, 3, 3, 3, 3, 2), (5, 2, 2, 2, 2, 2) for 6 layers, (10, 6, 3, 3, 3), (5, 4, 2, 2, 2) for 5 layers, (10, 6, 6, 3), (5, 4, 4, 2) for 4 layers, (20, 6, 6), (10, 4, 4) for 3 layers, (32, 20), (16, 10) for 2 layers and (320), (160) for 1 layer.
+While a larger inner dimension leads to a better performance, no major effect can be observed regarding the number of layers.
+The final projection additionally improves the performance, see \refTab{table:features_w2v_proj}.
+We can additionally vary the kernel size and final norm.
+
+\subsubsection{Inner Dimensions}
+As a larger inner dimension clearly showed better performance for the \wvtwo \gls{FE}, we increase the inner dimensions of the i6 features as well.
+However, as shown in \refTab{table:features_i6_size}, neither increasing the inner dimension of the first layer nor the second layer helps to improve the performance.
+
+\subsubsection{First Layer Window Size}
+What role does the size of the first layer play? Is it important to cover one pitch period (max. fundamental period)?
+By default, I use a window size of \SI{16}{\milli\second}.
+Experiments on using different window sizes in \refTab{table:features_window_size} suggest that a larger window size has a detrimental effect while going from the default \SI{16}{\milli\second} down to \SI{10}{\milli\second} shows a marginal improvement.
 
 \section{Results}
+The results for the different feature extractors used with a competitive \gls{CTC} model on Librispeech are shown in \refTab{table:features_general}.
+\input{tables/features_general}
+
+\subsection{Model Sizes}
+Results are in the following tables.
+\input{tables/features_w2v_size}
+\input{tables/features_w2v_proj}
+\input{tables/features_i6_size}
+\input{tables/features_window_size}
+
+\subsection{\wvtwo Pre-Training}
+\input{tables/features_pretraining}
 

### 2023-02-27 main.tex
add draft abstract

@@
 \begin{abstract}
+Raw waveform modeling is an interesting research topic for \gls{ASR}.
+\wvtwo has seen large success recently and makes use of a convolutional \gls{FE} which operates directly on the speech waveform.
+Typically, this \gls{FE} is seen as part of the overall architecture, but this work investigates this component specifically.
+It is used as a replacement for classical feature extraction methods in a \gls{CTC} setup and compared to an alternative neural \gls{FE} approach.
+Finally, we analyze the learned representations of both neural \glspl{FE}.
 \end{abstract}

### 2023-02-27 main.tex
FE dedicated command

--- 2023-02-26 main.tex
@@
 \begin{abstract}
+Raw waveform modeling is an interesting research topic for \gls{ASR}.
+\wvtwo has seen large success recently and makes use of a convolutional \fe which operates directly on the speech waveform.
+Typically, this \fe is seen as part of the overall architecture, but this work investigates this component specifically.
+It is used as a replacement for classical feature extraction methods in a \gls{CTC} setup and compared to an alternative neural \fe approach.
+Finally, we analyze the learned representations of both neural \fes.
 \end{abstract}

### 2023-02-27 main.tex
add draft notes for introduction paragraphs
add draft notes for related work

@@
 \section{Introduction}
+\draft{
+Why is raw waveform modeling interesting?
+
+\wvtwo is very successful and uses neural \fe.
+
+Comparable to other neural feature extraction methods, but not closely studied in the literature.
+Maybe some references or name some works which are similar.
+
+Highlight contributions of this paper.
+}
 
 \section{Related work}
+\draft{
+Works on raw waveform acoustic modeling.
+
+Different \wv papers \cite{facebook2019wav2vec, facebook2019vqwav2vec, facebook2020wav2vec2, facebook2020xlsr}.
+Describe differences between the approaches.
+Besides modifications of the self-supervised training criteria -- moving from the future time step prediction in \cite{facebook2019wav2vec} to masked time step prediction -- and incorporating quantization modules, also the architecture has been revised.
+Fully convolutional vs. \transformer.
+Differences in \fe.
+
+Think about how to properly reference \cite{vieting2021waveform} without disclosing author information.
+}
 

### 2023-02-27 main.tex
add todos/questions at end
update related work section, add w2v fe details

@@
 Works on raw waveform acoustic modeling.
+Think about how to properly reference \cite{vieting2021waveform} without disclosing author information.
 
@@
 
-Think about how to properly reference \cite{vieting2021waveform} without disclosing author information.
+\cite{facebook2019wav2vec} uses 5 layers with kernel sizes (10, 8, 4, 4, 4) and strides which are always exactly half the kernel size.
+All have 512 channels, a group normalization layer with a single group and a \relu nonlinearity.
+\cite{facebook2020wav2vec2} uses seven layers with kernel sizes (10, 3, 3, 3, 3, 2, 2), strides (5, 2, 2, 2, 2, 2, 2), and a GELU activation function.
+The first layer applies group normalization and finally, layer normalization and a linear projection is added.
 }
@@
 
+\section{To Dos/Questions}
+\begin{itemize}
+  \item \wvtwo feature extractor vs. feature encoder
+\end{itemize}
+
 \bibliographystyle{IEEEtran}

### 2023-02-27 main.tex
add text for feature extractors

@@
 \section{Methods}
+We always normalize the waveform to zero mean and unit variance.
+Then, it is input to a \fe.
+The \fe is followed by the remaining \gls{AM}.
+While the separation between \fe and remaining \gls{AM} is not sharp for neural feature extraction, we always refer to the part that replaces the traditional feature extraction as \fe.
+
 \subsection{Feature Extractors}
-In this work, we compare different feature extraction methods -- namely log Mel filterbank features, Gammatone features \cite{schlueter:icassp07}, \gls{SCF} features \cite{tuske2018:waveform} and the \gls{FE} of the \wvtwo model \cite{facebook2020wav2vec2}.
+\subsubsection{Standard Feature Extraction}
+As a reference, we use standard, hand-designed \fes -- namely log Mel filterbank features and Gammatone features.
+To compute the log Mel filterbank features, first the \gls{STFT} of the waveform is computed with a window size of \SI{25}{\milli\second} and a shift of \SI{10}{\milli\second}.
+We keep the square of the magnitude and perform Mel warping to obtain a 80-dim vector.
+Finally, $\log_{10}$ and normalization are applied.
 
+Gammatone features apply a gammatone filterbank to the pre-emphasized speech signal, perform temporal integration over each channel using a low pass filter, i.e., the Hanning window with a size of \SI{25}{\milli\second} and a shift of \SI{10}{\milli\second}, compress using the $10^{th}$ root and finally compute the \gls{DCT} of the values \cite{schlueter:icassp07}.
+Further, the the resulting 50-dim features are normalized.
+
+\subsubsection{\wvtwo Feature Extractor}
+The \gls{FE} of the \wvtwo model \cite{facebook2020wav2vec2}.
+
+\subsubsection{Supervised Convolutional Features}
+As an alternative comparable neural feature extraction method, we use \gls{SCF} \cite{tuske2018:waveform}.
+Similarly, a convolutional filterbank with learnable parameters is applied to the waveform.
+As in the case of Gammatones, a temporal integration is performed over each channel.
+However, in this case multiple filters are used for temporal integration allowing for multi-resolutional processing.
+Additionally, these filters are learned during training.
+We set the size of the first filterbank to 256, its strides to 10 and the number of channels to 150.
+The second convolutional filter applies 5 temporal integration filters with a size of 40 and strides of 16 each.
+Since the output of all 5 filters is stacked, we have a resulting feature dimension of 750.
+
+\section{Experiments}
 \subsection{Model Sizes}

### 2023-03-01 main.tex
w2v fe and \ms command

@@
 \cite{facebook2020wav2vec2} uses seven layers with kernel sizes (10, 3, 3, 3, 3, 2, 2), strides (5, 2, 2, 2, 2, 2, 2), and a GELU activation function.
-The first layer applies group normalization and finally, layer normalization and a linear projection is added.
+The first layer applies group normalization and finally, layer normalization and a linear projection are added.
 }
@@
 As a reference, we use standard, hand-designed \fes -- namely log Mel filterbank features and Gammatone features.
-To compute the log Mel filterbank features, first the \gls{STFT} of the waveform is computed with a window size of \SI{25}{\milli\second} and a shift of \SI{10}{\milli\second}.
+To compute the log Mel filterbank features, first the \gls{STFT} of the waveform is computed with a window size of \ms{25} and a shift of \ms{10}.
 We keep the square of the magnitude and perform Mel warping to obtain a 80-dim vector.
@@
 
-Gammatone features apply a gammatone filterbank to the pre-emphasized speech signal, perform temporal integration over each channel using a low pass filter, i.e., the Hanning window with a size of \SI{25}{\milli\second} and a shift of \SI{10}{\milli\second}, compress using the $10^{th}$ root and finally compute the \gls{DCT} of the values \cite{schlueter:icassp07}.
+Gammatone features apply a gammatone filterbank to the pre-emphasized speech signal, perform temporal integration over each channel using a low pass filter, i.e., the Hanning window with a size of \ms{25} and a shift of \ms{10}, compress using the $10^{th}$ root and finally compute the \gls{DCT} of the values \cite{schlueter:icassp07}.
 Further, the the resulting 50-dim features are normalized.
@@
 \subsubsection{\wvtwo Feature Extractor}
-The \gls{FE} of the \wvtwo model \cite{facebook2020wav2vec2}.
+The \gls{FE} of the \wvtwo model \cite{facebook2020wav2vec2} mainly consists of 7 convolutional layers.
+They are configured with kernel sizes (10, 3, 3, 3, 3, 2, 2), strides (5, 2, 2, 2, 2, 2, 2), and a GELU activation function.
+The first layer applies group normalization and finally, layer normalization and a linear projection is added.
+The total receptive field of the \wvtwo \fe is \ms{9999} and the shift between consecutive frames is \ms{20}.
+Since the other \fes operate at a shift of \ms{10} and the \wvtwo \fe is built in a modular way, we remove the last layer with a stride of 2 in order to achieve features at the same frame rate.
+This reduces the total receptive field to \ms{9999}.
 
@@
 Since the output of all 5 filters is stacked, we have a resulting feature dimension of 750.
+The total receptive field of the \gls{SCF} is \ms{9999}.
 
@@
 What role does the size of the first layer play? Is it important to cover one pitch period (max. fundamental period)?
-By default, I use a window size of \SI{16}{\milli\second}.
-Experiments on using different window sizes in \refTab{table:features_window_size} suggest that a larger window size has a detrimental effect while going from the default \SI{16}{\milli\second} down to \SI{10}{\milli\second} shows a marginal improvement.
+By default, I use a window size of \ms{16}.
+Experiments on using different window sizes in \refTab{table:features_window_size} suggest that a larger window size has a detrimental effect while going from the default \ms{16} down to \ms{10} shows a marginal improvement.
 

### 2023-03-02 figures/first_layer.tex
add figures for first layer impulse response

\begin{figure*}[htb]
	% blog post on how to use pgf from matplotlib to latex
	% https://timodenk.com/blog/exporting-matplotlib-plots-to-latex/
	% \usepackage{layouts} and \printinunitsof{cm}\prntlen{\textwidth} show that textwidth is 17cm here.
	\centering
	\input{figures/first_layer.pgf}
	\caption{First layer impulse responses.}
	\label{fig:first_layer}
\end{figure*}

### 2023-03-02 main.tex
to dos in red
add figures for first layer impulse response
add text to introduction

@@
 \usepackage{multirow}
+\usepackage{pgfplots}
 \usepackage{siunitx}
@@
 \section{Introduction}
+\glspl{AM} for \gls{ASR} traditionally makes use of hand-designed feature extraction methods like log Mel filterbank features \addref or Gammatone features \cite{schlueter:icassp07}.
+These techniques are motivated by insights into the properties of the human auditory system.
+However, their fixed design inevitably leads to information loss during the feature extraction.
+
+This issue can be addressed by neural \fes.
+Since their parameters are learned during training, these \fes can adapt to the need of \glspl{AM} and the loss of particularly helpful information may be avoided.
+A number of works have been proposed in this direction, often using convolutional \glspl{NN} \addref but also applying parametrized functions \cite{ravanelli2018sincnet} and other architectures \cite{sainath2015cldnn}.
+
+Recently, the \wvtwo framework \cite{facebook2020wav2vec2} has received large attention.
+It is mainly known for its contrastive self-supervised pre-training, which allows to pre-train a model on large amounts of unlabeled audio data.
+Subsequently, it can be fine-tuned for a specific downstream task at hand.
+While the context network with multiple \transformer blocks constitutes the major part of the architecture, the \wvtwo model also makes use of a convolutional feature encoder.
+\TODO{feature encoder vs. \fe here.}
+
+While the \wvtwo \fe is similar to other neural feature extraction methods, it is not yet closely studied in the literature.
+Choi et al. study the feature encoder\footnote{K. Choi and E. J. Yeo, “Opening the black box of wav2vec feature encoder,” arXiv preprint arXiv:2210.15386, 2022.}.%\cite{choi2022w2v2fe}
+
 \draft{
-Why is raw waveform modeling interesting?
-
-\wvtwo is very successful and uses neural \fe.
-
 Comparable to other neural feature extraction methods, but not closely studied in the literature.
@@
 
+\subsection{Frequency Response of Learned Filters}
+\input{figures/first_layer}
+
 \section{Conclusions}
@@
 
+{
+\color{red}
 \section{To Dos/Questions}
@@
 \end{itemize}
+}
 

### 2023-03-03 main.tex
example of arxiv citation in footnote
update introduction/related work

@@
 % imports
+\usepackage{bibentry}
 \usepackage[acronym,toc,shortcuts,nonumberlist]{glossaries}
@@
 \usepackage{xcolor}
+
+\nobibliography*  % used together with bibentry package to allow footnote bibentry citations
 
@@
 While the \wvtwo \fe is similar to other neural feature extraction methods, it is not yet closely studied in the literature.
-Choi et al. study the feature encoder\footnote{K. Choi and E. J. Yeo, “Opening the black box of wav2vec feature encoder,” arXiv preprint arXiv:2210.15386, 2022.}.%\cite{choi2022w2v2fe}
+\cite{choi2022w2v2fe} studies the feature encoder by feeding synthesized sine waves.
+The authors observe that enough temporal detail is obtained and information about fundamental frequencies is represented in the model's output.
+In particular, they claim that the frequencies are distributed linearly over the Hertz scale, unlike perceptually motivated scales like the Mel scale or the Greenwood function for Gammatone features.
+\draft{
+Other works that analyze \wvtwo models rather focus on the large \transformer part of the network which operates at a higher level of abstraction \cite{livescu2021wav2vec_analysis}\TODO{updated version and add more papers (references in Choi et al.
+or Livescu et al.}.
+}
 
 \draft{
-Comparable to other neural feature extraction methods, but not closely studied in the literature.
-Maybe some references or name some works which are similar.
-
 Highlight contributions of this paper.
@@
 The first layer applies group normalization and finally, layer normalization and a linear projection are added.
+
+\cite{asapp2022performance} investigates different settings for the \wvtwo \fe and shows that the required computation can be significantly reduced while maintaining \TODO{similar} performance.
+However, this is always done in the whole \wvtwo framework and not \TODO{what is different to us?}
 }

### 2023-03-03 main.tex
add text, mainly experiments

@@
 \section{Methods}
-We always normalize the waveform to zero mean and unit variance.
-Then, it is input to a \fe.
+This work compares different \fes' applicability for \gls{ASR}.
+We always normalize the waveform to zero mean and unit variance before it is input to a \fe.
 The \fe is followed by the remaining \gls{AM}.
@@
 Additionally, these filters are learned during training.
-We set the size of the first filterbank to 256, its strides to 10 and the number of channels to 150.
+By default, we set the size of the first filterbank to 256, its strides to 10 and the number of channels to 150.
 The second convolutional filter applies 5 temporal integration filters with a size of 40 and strides of 16 each.
@@
 
-\section{Experiments}
+\subsection{Acoustic Model}
+The \gls{AM} mainly consists of \gls{VGG} \addref blocks for downsampling and 12 \conformer blocks.
+The \gls{VGG} downsampling uses three 3x3-convolutional layers with 32, 64 and 64 channels, respectively.
+In total, a subsampling factor of 4 in the time dimension is achieved by the strided convolutions so that the \conformer operates on frames with a shift of \SI{40}{\milli\second}.
+The feature dimension is reduced by a factor of 2 due to max pooling and then multiplied by the number of channels in the last convolution resulting in a total increase by factor 32 in our case.
+\draft{
+There is a linear layer and dropout here.
+The \conformer blocks should be described here.
+}
+The configuration follows the setup in \cite{zeineldeen2022robustconformer} \TODO{check!}.
+
+\subsection{Connectionist Temporal Classification}
+We use a \gls{CTC} model for our experiments.
+It follows the setup described for the \gls{CTC} model in \cite{zhou2022efficient}.
+\TODO{Cite Wei's paper and add more description of the CTC model.}
+All models are trained in a purely supervised fashion unless explicitly stated otherwise.
+\TODO{Do we use an LM?}
+
+\section{Experimental Setup}
+\subsection{Data}
+We conduct experiments on the 960h LibriSpeech corpus \cite{panayotov2015librispeech}.
+For the experiments using a pre-trained \wvtwo \fe, the same 960h were used in pre-training.
+\draft{The \gls{LM} was trained on the default LibriSpeech \gls{LM} training data which includes the transcriptions of the training set along with the additional
+800M-word text-only data.}
+The evaluation is done on the standard \devclean, \devother, \testclean and \testother sets.
+
+\subsection{Training Details}
+\draft{We follow the setup in \TODO{Wei's paper} \addref.}
+The batch size is 640k samples, which corresponds to \SI{40}{\second} of speech waveform and the gradients are accumulated over 3 steps.
+The \TODO{NAdam} optimizer uses a \gls{OCLR} schedule with a peak learning rate of \TODO{X} as described in .
+Each model is trained on a single \gls{GPU}.
+\TODO{What about chunking? More details needed?}
+
+\section{Results}
 \subsection{Model Sizes}
@@
   \item \wvtwo feature extractor vs. feature encoder
+  \item check for arxiv citations
 \end{itemize}

### 2023-03-03 main.tex
update acknowledgements
merge two results sections

@@
 \section{Results}
+The results for the different feature extractors used with a competitive \gls{CTC} model on Librispeech are shown in \refTab{table:features_general}.
+\input{tables/features_general}
+
 \subsection{Model Sizes}
@@
 Can we improve the performance by making the model larger?
+
 \subsubsection{wav2vec 2.0 Components}
@@
 We can additionally vary the kernel size and final norm.
+\input{tables/features_w2v_size}
+\input{tables/features_w2v_proj}
 
@@
 However, as shown in \refTab{table:features_i6_size}, neither increasing the inner dimension of the first layer nor the second layer helps to improve the performance.
+\input{tables/features_i6_size}
 
@@
 Experiments on using different window sizes in \refTab{table:features_window_size} suggest that a larger window size has a detrimental effect while going from the default \ms{16} down to \ms{10} shows a marginal improvement.
-
-\section{Results}
-The results for the different feature extractors used with a competitive \gls{CTC} model on Librispeech are shown in \refTab{table:features_general}.
-\input{tables/features_general}
-
-\subsection{Model Sizes}
-Results are in the following tables.
-\input{tables/features_w2v_size}
-\input{tables/features_w2v_proj}
-\input{tables/features_i6_size}
 \input{tables/features_window_size}
@@
 \ifinterspeechfinal
-     Personal information revealing author identity
+     The authors would like to thank Wei Zhou for providing the baseline \gls{CTC} model.
 \else
-     Anonymized information not revealing author identity
+     The acknowledgements would likely reveal the author identity and are therefore hidden for the double-blind review process.
 \fi

### 2023-03-05 main.tex
text, mainly results

@@
 \subsubsection{Supervised Convolutional Features}
-As an alternative comparable neural feature extraction method, we use \gls{SCF} \cite{tuske2018:waveform}.
+As an alternative comparable neural feature extraction method, we use \gls{SC} features \cite{tuske2018:waveform, vieting2021waveform}.
 Similarly, a convolutional filterbank with learnable parameters is applied to the waveform.
@@
 Since the output of all 5 filters is stacked, we have a resulting feature dimension of 750.
-The total receptive field of the \gls{SCF} is \ms{9999}.
+The total receptive field of the \gls{SC} features is \ms{9999}.
 
@@
 \subsection{Data}
-We conduct experiments on the 960h LibriSpeech corpus \cite{panayotov2015librispeech}.
+We conduct experiments on the 960h LibriSpeech corpus \cite{panayotov2015librispeech}, which consists of read English speech.
 For the experiments using a pre-trained \wvtwo \fe, the same 960h were used in pre-training.
@@
 \section{Results}
-The results for the different feature extractors used with a competitive \gls{CTC} model on Librispeech are shown in \refTab{table:features_general}.
+The results, namely the \glspl{WER} on the LibriSpeech \textit{dev} and \textit{test} sets, for the different \fes are shown in \refTab{table:features_general}.
+We can observe that all are within a close range.
+The \gls{SC} features slightly outperform the Gammatone baseline, while log Mel features are again marginally better and the \wvtwo \fe yields the best results.
+This demonstrates that learnable neural \fes are competitive with hand-designed methods in a purely supervised setup with 960h of training data.
+However, a clear advantage of these techniques cannot be deduced from the experimental results.
 \input{tables/features_general}
@@
 \subsection{Model Sizes}
-By default, I use 40k parameters for i6 features compared to 4M for \wvtwo.
-Can we improve the performance by making the model larger?
+It is striking that the \wv2 \fe uses 100x more parameters than the \gls{SC} architecture.
+We therefore study the contribution of its components in \refSec{sec:w2v_components} and increase the number of parameters in the \gls{SC} model in \refSec{sec:scf_size}.
+Additionally, the size of the first window operating on the waveform is analyzed in \refSec{sec:scf_first_window}.
 
-\subsubsection{wav2vec 2.0 Components}
+\subsubsection{Dissecting the wav2vec 2.0 Feature Extractor}
+\label{sec:w2v_components}
 To understand where \wvtwo uses the parameters and how much they contribute to the \gls{FE}'s performance, we run it with different configurations.
 The results for different widths and depths are shown in \refTab{table:features_w2v_size}.
-% Kernel sizes and strides were chosen as (10, 3, 3, 3, 3, 2), (5, 2, 2, 2, 2, 2) for 6 layers, (10, 6, 3, 3, 3), (5, 4, 2, 2, 2) for 5 layers, (10, 6, 6, 3), (5, 4, 4, 2) for 4 layers, (20, 6, 6), (10, 4, 4) for 3 layers, (32, 20), (16, 10) for 2 layers and (320), (160) for 1 layer.
-While a larger inner dimension leads to a better performance, no major effect can be observed regarding the number of layers.
-The final projection additionally improves the performance, see \refTab{table:features_w2v_proj}.
-We can additionally vary the kernel size and final norm.
+The kernel sizes and strides were chosen as (10, 3, 3, 3, 3, 2), (5, 2, 2, 2, 2, 2) for 6 layers, (10, 6, 3, 3, 3), (5, 4, 2, 2, 2) for 5 layers, (10, 6, 6, 3), (5, 4, 4, 2) for 4 layers, (20, 6, 6), (10, 4, 4) for 3 layers and (32, 20), (16, 10) for 2 layers.
+The results show that a larger inner dimension leads to a better performance
+In contrast, no major effect can be observed regarding the number of layers.
+We also tried a variant with only one layer using kernel size 320 and stride 160 which did not converge.
+The final projection, which comprises about \TODO{400k} parameters, additionally improves the performance as shown in \refTab{table:features_w2v_proj}.
 \input{tables/features_w2v_size}
@@
 
-\subsubsection{Inner Dimensions}
-As a larger inner dimension clearly showed better performance for the \wvtwo \gls{FE}, we increase the inner dimensions of the i6 features as well.
-However, as shown in \refTab{table:features_i6_size}, neither increasing the inner dimension of the first layer nor the second layer helps to improve the performance.
-\input{tables/features_i6_size}
+\subsubsection{\draft{Inner Dimensions}}
+\label{sec:scf_size}
+As a larger inner dimension clearly showed better performance for the \wvtwo \gls{FE}, we increase the inner dimensions of the \gls{SC} features as well.
+However, as shown in \refTab{table:features_scf_size}, neither increasing the number of filters operating on the waveform nor the number of kernels for the multi-resolutional temporal integration helps to improve the performance.
+Note that the number of parameters of the full model is even larger than for the \wvtwo model in \refTab{table:features_general}, however, a large portion of those parameters is consumed by the linear projection of the features after the \gls{VGG} block.
+\input{tables/features_scf_size}
 
 \subsubsection{First Layer Window Size}
+\label{sec:scf_first_window}
+\draft{
 What role does the size of the first layer play? Is it important to cover one pitch period (max. fundamental period)?
@@
 Experiments on using different window sizes in \refTab{table:features_window_size} suggest that a larger window size has a detrimental effect while going from the default \ms{16} down to \ms{10} shows a marginal improvement.
+}
+Note that all experiments in \refSec{sec:scf_size} use a window size of \ms{16}.
 \input{tables/features_window_size}

### 2023-03-05 tables/create_tables.sh
update tables

@@ -10,7 +10,7 @@ fi
 python ${EXCEL2LATEX} --file_in features.xlsx --all_lines --vertical_merge --sheet general --wer_cols E
 python ${EXCEL2LATEX} --file_in features.xlsx --all_lines --vertical_merge --sheet pretraining --wer_cols C
 python ${EXCEL2LATEX} --file_in features.xlsx --all_lines --vertical_merge --sheet window_size --wer_cols C
-python ${EXCEL2LATEX} --file_in features.xlsx --all_lines --vertical_merge --sheet i6_size --wer_cols E
+python ${EXCEL2LATEX} --file_in features.xlsx --all_lines --vertical_merge --sheet scf_size --wer_cols E
 python ${EXCEL2LATEX} --file_in features.xlsx --all_lines --vertical_merge --sheet w2v_size --wer_cols D
 python ${EXCEL2LATEX} --file_in features.xlsx --all_lines --vertical_merge --sheet w2v_proj --wer_cols C
 

### 2023-03-05 tables/create_tables.sh
tables scf

@@ -10,7 +10,7 @@ fi
 python ${EXCEL2LATEX} --file_in features.xlsx --all_lines --vertical_merge --sheet general --wer_cols E
 python ${EXCEL2LATEX} --file_in features.xlsx --all_lines --vertical_merge --sheet pretraining --wer_cols C
 python ${EXCEL2LATEX} --file_in features.xlsx --all_lines --vertical_merge --sheet window_size --wer_cols C
-python ${EXCEL2LATEX} --file_in features.xlsx --all_lines --vertical_merge --sheet scf_size --wer_cols E
+python ${EXCEL2LATEX} --file_in features.xlsx --vertical_merge --sheet scf_size --wer_cols E
 python ${EXCEL2LATEX} --file_in features.xlsx --all_lines --vertical_merge --sheet w2v_size --wer_cols D
 python ${EXCEL2LATEX} --file_in features.xlsx --all_lines --vertical_merge --sheet w2v_proj --wer_cols C
 

### 2023-03-05 main.tex
text, mainly results

@@
 % imports
-\usepackage{bibentry}
 \usepackage[acronym,toc,shortcuts,nonumberlist]{glossaries}
@@
 \usepackage{siunitx}
+\usepackage{url}
 \usepackage{xcolor}
-
-\nobibliography*  % used together with bibentry package to allow footnote bibentry citations
 
@@
 \label{sec:scf_first_window}
-\draft{
-What role does the size of the first layer play? Is it important to cover one pitch period (max. fundamental period)?
-By default, I use a window size of \ms{16}.
-Experiments on using different window sizes in \refTab{table:features_window_size} suggest that a larger window size has a detrimental effect while going from the default \ms{16} down to \ms{10} shows a marginal improvement.
-}
+Furthermore, the effect of the first layer's window size is studied and the results are shown in \refTab{table:features_window_size}.
+They suggest that smaller window sizes than in \cite{tuske2018:waveform} perform better in the context of a \conformer \gls{CTC} model.
+In particular, it does not seem to be necessary to cover one full \draft{pitch period (max. fundamental period) \TODO{check detailed explanation} which is around \ms{16}}.
+Instead, the best result is achieved with a window size of \ms{10}.
 Note that all experiments in \refSec{sec:scf_size} use a window size of \ms{16}.
@@
 \subsection{\wvtwo Pre-Training}
+The main appeal of the \wvtwo framework is the ability to pre-train the model on unlabeled audio data.
+\cite{facebook2020wav2vec2} even reports relative \gls{WER} improvements of about 10\% on the \textit{other} subsets for their experiments on LibriSpeech when pre-training on the same data as used in supervised training compared to pure supervised training from scratch.
+Here, we study this effect for the \fe only by using the parameters of an \fe pre-trained on LibriSpeech\footnote{\raggedright\url{https://github.com/facebookresearch/fairseq/blob/main/examples/wav2vec/README.md}}, so no external data is added.
+During the supervised training, the \fe weights can be further trained or kept frozen.
+As shown in \refTab{table:features_pretraining}, no positive effect can be observed here.
+In fact, using a frozen pre-trained \fe shows clearly worse performance than training the \fe purely supervised from scratch.
+We hypothesize that the pre-training mainly benefits the \transformer encoder and not the \fe while the mismatch between the \transformer encoder in pre-training and \conformer encoder in our setup could also pose a challenge for the case that the \fe is frozen.
 \input{tables/features_pretraining}

### 2023-03-05 tables/features_pretraining.tex
update tables
tables: remove unused and add double lines

@@
 \centering
-\caption{Effect of pre-training the wav2vec 2.0 feature extractor using the unsupervised loss on the same data (Librispeech 960h)}
+\caption{Effect of pre-training the wav2vec 2.0 feature extractor using the unsupervised loss on the same data (Librispeech 960h).}
 \label{table:features_pretraining}
@@
 \multicolumn{2}{|c|}{Feature Extractor} &       {WER} \\\hline
-                              Trainable & Pretrained & {dev-other} \\\hline
+                              Trainable & Pretrained & {dev-other} \\\hline\hline
                                     yes &         no &         6.8 \\\cline{2-3}

### 2023-03-05 tables/create_tables.sh
general table: WERs for all subsets template

@@ -7,7 +7,7 @@ else
     EXCEL2LATEX=/u/vieting/Documents/text/excel-to-latex/excel_to_latex/excel_to_latex.py
 fi
 
-python ${EXCEL2LATEX} --file_in features.xlsx --all_lines --vertical_merge --sheet general --wer_cols E
+python ${EXCEL2LATEX} --file_in features.xlsx --all_lines --vertical_merge --sheet general --wer_cols E F G H
 python ${EXCEL2LATEX} --file_in features.xlsx --all_lines --vertical_merge --sheet pretraining --wer_cols C
 python ${EXCEL2LATEX} --file_in features.xlsx --all_lines --vertical_merge --sheet window_size --wer_cols C
 python ${EXCEL2LATEX} --file_in features.xlsx --vertical_merge --sheet scf_size --wer_cols E

### 2023-03-05 tables/features_pretraining.tex
tables: WERs for all subsets template

@@
 \label{table:features_pretraining}
-\begin{tabular}{|c|c|S|}
+\begin{tabular}{|c|c|S|c|S|c|}
 \hline
-\multicolumn{2}{|c|}{Feature Extractor} &       {WER} \\\hline
-                              Trainable & Pretrained & {dev-other} \\\hline\hline
-                                    yes &         no &         6.8 \\\cline{2-3}
-                                        &        yes &         6.9 \\\cline{1-1}\cline{3-3}
-                                     no &            &         7.1 \\
+\multicolumn{2}{|c|}{Feature Extractor} & \multicolumn{4}{c|}{{WER}} \\\cline{3-6}
+                 \multicolumn{2}{|c|}{} & \multicolumn{2}{c|}{{dev}} & \multicolumn{2}{c|}{{test}} \\\hline
+                              Trainable & Pretrained &                    {clean} & other &                     {clean} & other \\\hline\hline
+                                    yes &         no &                            &   6.8 &                             &       \\\cline{2-6}
+                                        &        yes &                            &   6.9 &                             &       \\\cline{1-1}\cline{3-6}
+                                     no &            &                            &   7.1 &                             &       \\
 \hline

### 2023-03-05 tables/create_tables.sh

@@ -8,9 +8,9 @@ else
 fi
 
 python ${EXCEL2LATEX} --file_in features.xlsx --all_lines --vertical_merge --sheet general --wer_cols E F G H
-python ${EXCEL2LATEX} --file_in features.xlsx --all_lines --vertical_merge --sheet pretraining --wer_cols C
-python ${EXCEL2LATEX} --file_in features.xlsx --all_lines --vertical_merge --sheet window_size --wer_cols C
-python ${EXCEL2LATEX} --file_in features.xlsx --vertical_merge --sheet scf_size --wer_cols E
-python ${EXCEL2LATEX} --file_in features.xlsx --all_lines --vertical_merge --sheet w2v_size --wer_cols D
-python ${EXCEL2LATEX} --file_in features.xlsx --all_lines --vertical_merge --sheet w2v_proj --wer_cols C
+python ${EXCEL2LATEX} --file_in features.xlsx --all_lines --vertical_merge --sheet pretraining --wer_cols C D E F
+python ${EXCEL2LATEX} --file_in features.xlsx --all_lines --vertical_merge --sheet window_size --wer_cols C D E F
+python ${EXCEL2LATEX} --file_in features.xlsx --vertical_merge --sheet scf_size --wer_cols E F G H
+python ${EXCEL2LATEX} --file_in features.xlsx --all_lines --vertical_merge --sheet w2v_size --wer_cols D E F G
+python ${EXCEL2LATEX} --file_in features.xlsx --all_lines --vertical_merge --sheet w2v_proj --wer_cols C D E F
 

### 2023-03-05 main.tex
general table: WERs for all subsets template
tables: WERs for all subsets template
inner dimension subsec header
add text for results of images

@@
 \usepackage{xcolor}
+
+\setlength{\tabcolsep}{2pt} % Narrower tables. Default value: 6pt
 
@@
 
-\subsubsection{\draft{Inner Dimensions}}
+\subsubsection{Larger Supervised Convolutional Architecture}
 \label{sec:scf_size}
@@
 \subsection{Frequency Response of Learned Filters}
+Finally, we analyze the learned filters of the neural \fes.
+This is done by plotting the frequency response of the filters in the first layer which operates on the raw waveform.
+As the order of learned filters in a convolutional layer is arbitrary, we sort the filters by the peak value of the frequency response.
+The plots are depicted in \refFig{fig:first_layer}.
+The color range is shared between all subplots.
+
+The well known distribution of the Gammatone filters is given in \refFig{fig:first_layer}(a).
+\refFig{fig:first_layer}(b) shows the frequency response of the learned \gls{SC} filters.
+In line with \cite{tuske2018:waveform}, bandpass filters with a non-linear distribution of center frequencies are learned.
+The distribution resembles the Gammatone filterbank and with higher center frequencies, the pass band is also wider.
+However, the stopband attenuation is clearly weaker.
+
+The distribution for the \wvtwo \fe's first layer in \refFig{fig:first_layer}(c) does not entail this degree of resemblance.
+Again, bandpass filters are learned, however, they have a wider passband, their center frequencies are distributed rather linearly and the stopband attenuation is weak.
+Interestingly, the frequency response of the pre-trained \wvtwo \fe in \refFig{fig:first_layer}(d) is very similar to \refFig{fig:first_layer}(c).
+This indicates that the properties learned in unsupervised pre-training and supervised \gls{CTC} training are similar.
 \input{figures/first_layer}

### 2023-03-05 main.tex
comment on improved GT baseline
nitpicking

@@
 \TODO{Cite Wei's paper and add more description of the CTC model.}
-All models are trained in a purely supervised fashion unless explicitly stated otherwise.
-\TODO{Do we use an LM?}
+All models are trained in a purely supervised fashion using the \gls{CTC} objective unless explicitly stated otherwise.
+A 4-gram \gls{LM} is used and integrated using shallow fusion \cite{gulcehre2015shallow} during decoding.
+\TODO{Check LM usage.}
 
@@
 The results, namely the \glspl{WER} on the LibriSpeech \textit{dev} and \textit{test} sets, for the different \fes are shown in \refTab{table:features_general}.
-We can observe that all are within a close range.
+The baseline performance using Gammatone features is improved compared to \cite{zhou2022efficient} by disabling gradient noise and re-tuning the peak learning rate.
+We can observe that all \fes are within a close range.
 The \gls{SC} features slightly outperform the Gammatone baseline, while log Mel features are again marginally better and the \wvtwo \fe yields the best results.
@@
 
-The well known distribution of the Gammatone filters is given in \refFig{fig:first_layer}(a).
-\refFig{fig:first_layer}(b) shows the frequency response of the learned \gls{SC} filters.
+The well known distribution of the Gammatone filters is given in \refFig{fig:first_layer}\,(a).
+\refFig{fig:first_layer}\,(b) shows the frequency response of the learned \gls{SC} filters.
 In line with \cite{tuske2018:waveform}, bandpass filters with a non-linear distribution of center frequencies are learned.
@@
 
-The distribution for the \wvtwo \fe's first layer in \refFig{fig:first_layer}(c) does not entail this degree of resemblance.
+The distribution for the \wvtwo \fe's first layer in \refFig{fig:first_layer}\,(c) does not entail this degree of resemblance.
 Again, bandpass filters are learned, however, they have a wider passband, their center frequencies are distributed rather linearly and the stopband attenuation is weak.
-Interestingly, the frequency response of the pre-trained \wvtwo \fe in \refFig{fig:first_layer}(d) is very similar to \refFig{fig:first_layer}(c).
+Interestingly, the frequency response of the pre-trained \wvtwo \fe in \refFig{fig:first_layer}\,(d) is very similar to \refFig{fig:first_layer}\,(c).
 This indicates that the properties learned in unsupervised pre-training and supervised \gls{CTC} training are similar.

### 2023-03-07 tables/features_pretraining.tex
tables: WER in percentage, update captions
table update caption
update results for other subsets

--- 2023-02-26 tables/features_pretraining.tex
@@
 \centering
-\caption{Effect of pre-training the wav2vec 2.0 feature extractor using the unsupervised loss on the same data (Librispeech 960h)}
+\caption{Effect of pre-training the wav2vec 2.0 feature extractor using the unsupervised loss on the same data (LibriSpeech 960h).}
 \label{table:features_pretraining}
-\begin{tabular}{|c|c|S|}
+\begin{tabular}{|c|c|S|c|S|c|}
 \hline
-\multicolumn{2}{|c|}{Feature Extractor} &       {WER} \\\hline
-                              Trainable & Pretrained & {dev-other} \\\hline
-                                    yes &         no &         6.8 \\\cline{2-3}
-                                        &        yes &         6.9 \\\cline{1-1}\cline{3-3}
-                                     no &            &         7.1 \\
+\multicolumn{2}{|c|}{Feature Extractor} & \multicolumn{4}{c|}{{WER [\%]}} \\\cline{3-6}
+                 \multicolumn{2}{|c|}{} &      \multicolumn{2}{c|}{{dev}} & \multicolumn{2}{c|}{{test}} \\\hline
+                              Trainable & Pretrained &                         {clean} & other &                     {clean} & other \\\hline\hline
+                                    yes &         no &                             2.9 &   6.8 &                             &   7.5 \\\cline{2-6}
+                                        &        yes &                             2.9 &   6.9 &                             &   7.5 \\\cline{1-1}\cline{3-6}
+                                     no &            &                             3.0 &   7.1 &                             &   7.8 \\
 \hline

### 2023-03-07 main.tex
move figure up
small updates, version sent to Ralf as draft

@@
 \draft{
-Other works that analyze \wvtwo models rather focus on the large \transformer part of the network which operates at a higher level of abstraction \cite{livescu2021wav2vec_analysis}\TODO{updated version and add more papers (references in Choi et al.
-or Livescu et al.}.
+Other works that analyze \wvtwo models rather focus on the large \transformer part of the network which operates at a higher level of abstraction \cite{livescu2021wav2vec_analysis}.
+\TODO{updated version and add more papers (references in Choi et al. or Livescu et al.}
 }
@@
 \cite{asapp2022performance} investigates different settings for the \wvtwo \fe and shows that the required computation can be significantly reduced while maintaining \TODO{similar} performance.
-However, this is always done in the whole \wvtwo framework and not \TODO{what is different to us?}
+However, this is always done in the whole \wvtwo framework.
+\TODO{Highlight differences to this paper.}
 }
@@
 
-Gammatone features apply a gammatone filterbank to the pre-emphasized speech signal, perform temporal integration over each channel using a low pass filter, i.e., the Hanning window with a size of \ms{25} and a shift of \ms{10}, compress using the $10^{th}$ root and finally compute the \gls{DCT} of the values \cite{schlueter:icassp07}.
+Gammatone features apply a Gammatone filterbank to the pre-emphasized speech signal, perform temporal integration over each channel using a low pass filter, i.e., the Hanning window with a size of \ms{25} and a shift of \ms{10}, compress using the $10^{th}$ root and finally compute the \gls{DCT} of the values \cite{schlueter:icassp07}.
 Further, the the resulting 50-dim features are normalized.
@@
 The first layer applies group normalization and finally, layer normalization and a linear projection is added.
-The total receptive field of the \wvtwo \fe is \ms{9999} and the shift between consecutive frames is \ms{20}.
+The total receptive field of the \wvtwo \fe is \ms{25} and the shift between consecutive frames is \ms{20}.
 Since the other \fes operate at a shift of \ms{10} and the \wvtwo \fe is built in a modular way, we remove the last layer with a stride of 2 in order to achieve features at the same frame rate.
-This reduces the total receptive field to \ms{9999}.
+This reduces the total receptive field to \ms{15}.
 
@@
 It follows the setup described for the \gls{CTC} model in \cite{zhou2022efficient}.
-\TODO{Cite Wei's paper and add more description of the CTC model.}
+\TODO{Add more description of the CTC model.}
 All models are trained in a purely supervised fashion using the \gls{CTC} objective unless explicitly stated otherwise.
-A 4-gram \gls{LM} is used and integrated using shallow fusion \cite{gulcehre2015shallow} during decoding.
-\TODO{Check LM usage.}
+The official LibriSpeech 4-gram \gls{LM} is used and integrated using shallow fusion \cite{gulcehre2015shallow} during decoding.
 
@@
 For the experiments using a pre-trained \wvtwo \fe, the same 960h were used in pre-training.
-\draft{The \gls{LM} was trained on the default LibriSpeech \gls{LM} training data which includes the transcriptions of the training set along with the additional
-800M-word text-only data.}
+The \gls{LM} was trained on the default LibriSpeech \gls{LM} training data which includes the transcriptions of the training set along with the additional 800M-word text only data.
 The evaluation is done on the standard \devclean, \devother, \testclean and \testother sets.
@@
 \subsection{Training Details}
-\draft{We follow the setup in \TODO{Wei's paper} \addref.}
+We generally follow the setup in \cite{zhou2022efficient}.
 The batch size is 640k samples, which corresponds to \SI{40}{\second} of speech waveform and the gradients are accumulated over 3 steps.
-The \TODO{NAdam} optimizer uses a \gls{OCLR} schedule with a peak learning rate of \TODO{X} as described in .
+The NAdam optimizer uses a \gls{OCLR} schedule with a peak learning rate of $3\cdot 10^{-4}$.
+In contrast to \cite{zhou2022efficient}, no gradient noise is applied.
 Each model is trained on a single \gls{GPU}.
-\TODO{What about chunking? More details needed?}
 
@@
 The results, namely the \glspl{WER} on the LibriSpeech \textit{dev} and \textit{test} sets, for the different \fes are shown in \refTab{table:features_general}.
-The baseline performance using Gammatone features is improved compared to \cite{zhou2022efficient} by disabling gradient noise and re-tuning the peak learning rate.
+The baseline performance using Gammatone features is improved compared to \cite{zhou2022efficient} mainly by disabling gradient noise and re-tuning the peak learning rate.
 We can observe that all \fes are within a close range.
@@
 \subsection{Model Sizes}
-It is striking that the \wv2 \fe uses 100x more parameters than the \gls{SC} architecture.
+It is striking that the \wv2 \fe uses two orders of magnitude more parameters than the \gls{SC} architecture.
 We therefore study the contribution of its components in \refSec{sec:w2v_components} and increase the number of parameters in the \gls{SC} model in \refSec{sec:scf_size}.
@@
 We also tried a variant with only one layer using kernel size 320 and stride 160 which did not converge.
-The final projection, which comprises about \TODO{400k} parameters, additionally improves the performance as shown in \refTab{table:features_w2v_proj}.
+The final projection, which comprises about 394k parameters, additionally improves the performance as shown in \refTab{table:features_w2v_proj}.
 \input{tables/features_w2v_size}
@@
 They suggest that smaller window sizes than in \cite{tuske2018:waveform} perform better in the context of a \conformer \gls{CTC} model.
-In particular, it does not seem to be necessary to cover one full \draft{pitch period (max. fundamental period) \TODO{check detailed explanation} which is around \ms{16}}.
+In particular, it does not seem to be necessary to cover one full \draft{pitch period (max. fundamental period) which is around \ms{16} for low voices}. %\TODO{check detailed explanation.}
 Instead, the best result is achieved with a window size of \ms{10}.
@@
 \subsection{Frequency Response of Learned Filters}
+\input{figures/first_layer}
 Finally, we analyze the learned filters of the neural \fes.
@@
 This indicates that the properties learned in unsupervised pre-training and supervised \gls{CTC} training are similar.
-\input{figures/first_layer}
 

### 2023-03-07 main.tex
update text, mainly introduction and related work

@@
 \end{abstract}
-\noindent\textbf{Index Terms}: speech recognition, feature extraction, raw waveform modeling
+\noindent\textbf{Index Terms}: speech recognition, feature extraction, raw waveform modeling, wav2vec 2.0
+
+% reset glossaries
+\glsresetall
 
 \section{Introduction}
-\glspl{AM} for \gls{ASR} traditionally makes use of hand-designed feature extraction methods like log Mel filterbank features \addref or Gammatone features \cite{schlueter:icassp07}.
+\glspl{AM} for \gls{ASR} traditionally makes use of hand-designed feature extraction methods like log Mel filterbank features or Gammatone features \cite{schlueter:icassp07}.
 These techniques are motivated by insights into the properties of the human auditory system.
@@
 Since their parameters are learned during training, these \fes can adapt to the need of \glspl{AM} and the loss of particularly helpful information may be avoided.
-A number of works have been proposed in this direction, often using convolutional \glspl{NN} \addref but also applying parametrized functions \cite{ravanelli2018sincnet} and other architectures \cite{sainath2015cldnn}.
+A number of works have been proposed in this direction, often using convolutional \glspl{NN} \cite{palaz2015convolutional, golik15:cnn, tuske2018:waveform} but also applying parametrized functions \cite{ravanelli2018sincnet} and other architectures \cite{sainath2015cldnn}.
 
@@
 \cite{choi2022w2v2fe} studies the feature encoder by feeding synthesized sine waves.
-The authors observe that enough temporal detail is obtained and information about fundamental frequencies is represented in the model's output.
-In particular, they claim that the frequencies are distributed linearly over the Hertz scale, unlike perceptually motivated scales like the Mel scale or the Greenwood function for Gammatone features.
-\draft{
-Other works that analyze \wvtwo models rather focus on the large \transformer part of the network which operates at a higher level of abstraction \cite{livescu2021wav2vec_analysis}.
-\TODO{updated version and add more papers (references in Choi et al. or Livescu et al.}
-}
+Other works that analyze \wvtwo models rather focus on the large \transformer part of the network which operates at a higher level of abstraction \cite{livescu2021wav2vec_analysis, fan21wvspeakerid, li2023exploration, dieck2022wav2vec}.
 
@@
 \section{Related work}
-\draft{
-Works on raw waveform acoustic modeling.
-Think about how to properly reference \cite{vieting2021waveform} without disclosing author information.
+This paper takes up the line of research that applies learnable feature extraction front-ends for \gls{ASR} \cite{palaz2015convolutional, golik15:cnn, tuske2018:waveform, ravanelli2018sincnet, sainath2015cldnn}.
+In particular, different neural \fes are compared to traditional feature extraction methods.
+We choose to investigate the supervised convolutional \fe from \cite{tuske2018:waveform}, that was also studied in \cite{vieting2021waveform}.
+Furthermore, the feature extractor of \wvtwo \cite{facebook2020wav2vec2} is used as a learnable front-end for supervised \gls{ASR} training.
 
-Different \wv papers \cite{facebook2019wav2vec, facebook2019vqwav2vec, facebook2020wav2vec2, facebook2020xlsr}.
-Describe differences between the approaches.
+The \wv framework has been established in a series of related papers \cite{facebook2019wav2vec, facebook2019vqwav2vec, facebook2020wav2vec2, facebook2020xlsr}.
 Besides modifications of the self-supervised training criteria -- moving from the future time step prediction in \cite{facebook2019wav2vec} to masked time step prediction -- and incorporating quantization modules, also the architecture has been revised.
-Fully convolutional vs. \transformer.
-Differences in \fe.
+While \cite{facebook2019wav2vec} uses a fully convolutional architecture, \cite{facebook2020wav2vec2} uses a large stack of \transformer blocks to contextualize the convolutional representations.
+%
+Also the \fe itself has been adjusted.
+%\cite{facebook2019wav2vec} uses 5 layers with kernel sizes (10, 8, 4, 4, 4) and strides which are always exactly half the kernel size.
+%All have 512 channels, a group normalization layer with a single group and a \relu nonlinearity.
+%\cite{facebook2020wav2vec2} uses seven layers with kernel sizes (10, 3, 3, 3, 3, 2, 2), strides (5, 2, 2, 2, 2, 2, 2), and a GELU activation function.
+%The first layer applies group normalization and finally, layer normalization and a linear projection are added.
+\cite{facebook2019wav2vec} uses 5 layers, a group normalization with a single group and a \relu nonlinearity.
+In contrast, 7 layers with group normalization in the first layer and a GELU activation function are used in \cite{facebook2020wav2vec2}.
+Additionally, layer normalization and a linear projection are added after the convolutional layers.
 
-\cite{facebook2019wav2vec} uses 5 layers with kernel sizes (10, 8, 4, 4, 4) and strides which are always exactly half the kernel size.
-All have 512 channels, a group normalization layer with a single group and a \relu nonlinearity.
-\cite{facebook2020wav2vec2} uses seven layers with kernel sizes (10, 3, 3, 3, 3, 2, 2), strides (5, 2, 2, 2, 2, 2, 2), and a GELU activation function.
-The first layer applies group normalization and finally, layer normalization and a linear projection are added.
+While most works analyzing the \wvtwo model focus on the \transformer part, there are papers which study the \fe as well.
+The authors of \cite{choi2022w2v2fe} feed synthesized sine waves to a pre-trained model and observe that enough temporal detail is obtained and information about fundamental frequencies is represented in the model's output.
+In particular, they claim that the frequencies are distributed linearly over the Hertz scale, unlike perceptually motivated scales like the Mel scale or the Greenwood function for Gammatone features.
+\cite{livescu2021wav2vec_analysis} states that the learned representations of the \fe are highly correlated with Mel spectrogram features as was determined by the \gls{CCA}, where the correlation is at a high plateau for \fe layers 3 to 7.
+This observation is supported by experimental evidence in \cite{dieck2022wav2vec}.
 
-\cite{asapp2022performance} investigates different settings for the \wvtwo \fe and shows that the required computation can be significantly reduced while maintaining \TODO{similar} performance.
-However, this is always done in the whole \wvtwo framework.
-\TODO{Highlight differences to this paper.}
-}
+\cite{asapp2022performance} investigates different settings for the \wvtwo \fe and shows that the required computation can be significantly reduced while maintaining similar performance.
+Furthermore, reducing the \fe's size in favor of a larger context network shows promising results.
+%In addition, a deeper variant with additional point-wise convolutional layers is proposed which outperforms its wider counterpart using higher inner dimensions.
+However, unlike our work this is always done in the whole \wvtwo framework including pre-training and fine-tuning and no insights into the learned representations are given.
 
@@
 They suggest that smaller window sizes than in \cite{tuske2018:waveform} perform better in the context of a \conformer \gls{CTC} model.
-In particular, it does not seem to be necessary to cover one full \draft{pitch period (max. fundamental period) which is around \ms{16} for low voices}. %\TODO{check detailed explanation.}
-Instead, the best result is achieved with a window size of \ms{10}.
+In particular, it does not seem to be necessary to cover one full pitch period which is around \ms{16} for low voices.
+Instead, the best result is achieved with a window size of \ms{10} and even smaller windows still achieve comparable performance.
 Note that all experiments in \refSec{sec:scf_size} use a window size of \ms{16}.

### 2023-03-07 tables/features_pretraining.tex
add more results to tables

@@
                               Trainable & Pretrained &                         {clean} & other &                     {clean} & other \\\hline\hline
-                                    yes &         no &                             2.9 &   6.8 &                             &   7.5 \\\cline{2-6}
-                                        &        yes &                             2.9 &   6.9 &                             &   7.5 \\\cline{1-1}\cline{3-6}
-                                     no &            &                             3.0 &   7.1 &                             &   7.8 \\
+                                    yes &         no &                             2.9 &   6.8 &                         3.3 &   7.5 \\\cline{2-6}
+                                        &        yes &                             2.9 &   6.9 &                         3.3 &   7.5 \\\cline{1-1}\cline{3-6}
+                                     no &            &                             3.0 &   7.1 &                         3.3 &   7.8 \\
 \hline

### 2023-03-07 main.tex
update AM text

@@
 The \gls{AM} mainly consists of \gls{VGG} \addref blocks for downsampling and 12 \conformer blocks.
+Its configuration mostly follows \cite{zeineldeen2022conformer} and \cite{zhou2022efficient}.
 The \gls{VGG} downsampling uses three 3x3-convolutional layers with 32, 64 and 64 channels, respectively.
@@
 The feature dimension is reduced by a factor of 2 due to max pooling and then multiplied by the number of channels in the last convolution resulting in a total increase by factor 32 in our case.
-\draft{
-There is a linear layer and dropout here.
-The \conformer blocks should be described here.
-}
-The configuration follows the setup in \cite{zeineldeen2022robustconformer} \TODO{check!}.
+After a linear projection to a 512-dimensional representation and dropout of 0.1, the 12 \conformer blocks with an inner dimension of 512, swapped convolution and multi-headed self-attention modules and 8 heads are applied.
 

### 2023-03-07 main.tex
move figure up

@@
 \subsection{\wvtwo Pre-Training}
+\input{figures/first_layer}
 The main appeal of the \wvtwo framework is the ability to pre-train the model on unlabeled audio data.
@@
 \subsection{Frequency Response of Learned Filters}
-\input{figures/first_layer}
 Finally, we analyze the learned filters of the neural \fes.

### 2023-03-08 main.tex
update text, add increasing w2v dims

@@
 \address{
-  $^1$RWTH Aachen University, Germany\\
+  $^1$Machine Learning and Human Language Technology,\\Computer Science Department, RWTH Aachen University, Germany\\
   $^2$AppTek GmbH, Germany}
@@
 Finally, we analyze the learned representations of both neural \fes.
+\TODO{Revise abstract and include results.}
 \end{abstract}
@@
 While the context network with multiple \transformer blocks constitutes the major part of the architecture, the \wvtwo model also makes use of a convolutional feature encoder.
-\TODO{feature encoder vs. \fe here.}
+Feature encoder \cite{facebook2020wav2vec2, facebook2020xlsr} and feature extractor \cite{asapp2022performance, vyas2022ondemand} are both common terms in the literature referring to the same part of the model and for consistency, we only use the term \acrfull{FE} here as it is more in line with other works on neural front-ends.
 
 While the \wvtwo \fe is similar to other neural feature extraction methods, it is not yet closely studied in the literature.
-\cite{choi2022w2v2fe} studies the feature encoder by feeding synthesized sine waves.
+\cite{choi2022w2v2fe} studies the \fe by feeding synthesized sine waves.
 Other works that analyze \wvtwo models rather focus on the large \transformer part of the network which operates at a higher level of abstraction \cite{livescu2021wav2vec_analysis, fan21wvspeakerid, li2023exploration, dieck2022wav2vec}.
@@
 Furthermore, reducing the \fe's size in favor of a larger context network shows promising results.
-%In addition, a deeper variant with additional point-wise convolutional layers is proposed which outperforms its wider counterpart using higher inner dimensions.
+In addition, a deeper variant with additional point-wise convolutional layers is proposed which outperforms its wider counterpart using higher inner dimensions.
 However, unlike our work this is always done in the whole \wvtwo framework including pre-training and fine-tuning and no insights into the learned representations are given.
@@
 The batch size is 640k samples, which corresponds to \SI{40}{\second} of speech waveform and the gradients are accumulated over 3 steps.
-The NAdam optimizer uses a \gls{OCLR} schedule with a peak learning rate of $3\cdot 10^{-4}$.
+The NAdam optimizer uses a \gls{OCLR} schedule with a peak learning rate of $3\cdot 10^{-4}$ and we train the model for 20 epochs.
 In contrast to \cite{zhou2022efficient}, no gradient noise is applied.
@@
 However, a clear advantage of these techniques cannot be deduced from the experimental results.
+\TODO{Add params for Mel filterbank and a note on linear layer using much more params for larger feature dim.}
 \input{tables/features_general}
@@
 \subsection{Model Sizes}
-It is striking that the \wv2 \fe uses two orders of magnitude more parameters than the \gls{SC} architecture.
+It is striking that the \wvtwo \fe uses two orders of magnitude more parameters than the \gls{SC} architecture.
 We therefore study the contribution of its components in \refSec{sec:w2v_components} and increase the number of parameters in the \gls{SC} model in \refSec{sec:scf_size}.
@@
 We also tried a variant with only one layer using kernel size 320 and stride 160 which did not converge.
-The final projection, which comprises about 394k parameters, additionally improves the performance as shown in \refTab{table:features_w2v_proj}.
+
+The proposed \fe in \cite{asapp2022performance} uses increasing inner dimensions and we use this approach in the last three lines of \refTab{table:features_w2v_size}.
+The first layer uses a dimension of 64/128 and it is doubled after each layer with an odd index.
+Additionally, we add point-wise convolutional layers after all but the first layer using the same inner dimension as the preceding layer for the last row of the table.
+It can be observed that these configurations perform similarly to the 6 layer 512-dim baseline while reducing the training time by about 21\%, 19\% and 14\%, respectively.
+
+The final projection, which comprises about 394k parameters, additionally improves the performance slightly on \devother while degrading minimally on the \textit{test} sets as shown in \refTab{table:features_w2v_proj}.
 \input{tables/features_w2v_size}
@@
 \begin{itemize}
-  \item \wvtwo feature extractor vs. feature encoder
   \item check for arxiv citations

### 2023-03-08 main.tex
update figure and corresponding text

@@
 \input{tables/features_w2v_proj}
+% has to be here to be shown on page 4
+\input{figures/first_layer}
 
@@
 \subsection{\wvtwo Pre-Training}
-\input{figures/first_layer}
 The main appeal of the \wvtwo framework is the ability to pre-train the model on unlabeled audio data.
@@
 Finally, we analyze the learned filters of the neural \fes.
-This is done by plotting the frequency response of the filters in the first layer which operates on the raw waveform.
+First, the frequency response of the filters in the first layer which operates on the raw waveform is plotted.
 As the order of learned filters in a convolutional layer is arbitrary, we sort the filters by the peak value of the frequency response.
-The plots are depicted in \refFig{fig:first_layer}.
-The color range is shared between all subplots.
+The plots are depicted in \refFig{fig:first_layer}\,(a)-(d), which share the same color range.
 
-The well known distribution of the Gammatone filters is given in \refFig{fig:first_layer}\,(a).
+The well known distribution of the Gammatone filters is given in \refFig{fig:first_layer}\,(a) as a reference.
 \refFig{fig:first_layer}\,(b) shows the frequency response of the learned \gls{SC} filters.
 In line with \cite{tuske2018:waveform}, bandpass filters with a non-linear distribution of center frequencies are learned.
+This confirms, that these filters are also preferred by more advanced neural back-ends like self-attention-based like \conformer which are superior in the exploitation of context information compared to simpler neural back-ends e.g.
+in \cite{tuske2014raw}.
 The distribution resembles the Gammatone filterbank and with higher center frequencies, the pass band is also wider.

### 2023-03-08 main.tex
small text updates, mel params in table 1

@@
 Additionally, these filters are learned during training.
-By default, we set the size of the first filterbank to 256, its strides to 10 and the number of channels to 150.
+By default, we set the size of the first filterbank to 160, its strides to 10 and the number of channels to 150.
 The second convolutional filter applies 5 temporal integration filters with a size of 40 and strides of 16 each.
 Since the output of all 5 filters is stacked, we have a resulting feature dimension of 750.
-The total receptive field of the \gls{SC} features is \ms{9999}.
+The total receptive field of the \gls{SC} features is \ms{40}.
 
@@
 However, a clear advantage of these techniques cannot be deduced from the experimental results.
-\TODO{Add params for Mel filterbank and a note on linear layer using much more params for larger feature dim.}
+The number of parameters for Gammatone and log Mel features refers to the \gls{FIR} and Mel filterbanks, respectively, that we use in our implementation even if they are not trainable.
+The differences in the total number of parameters are influenced by the size of the \fe, but also of the linear layer before the \conformer which grows large as the feature dimension increases.
 \input{tables/features_general}
@@
 This indicates that the properties learned in unsupervised pre-training and supervised \gls{CTC} training are similar.
+% for SCF, first 9 filters have peak at 0Hz, for wav2vec 75 filters
 

### 2023-03-09 main.tex
add text for first layer freq resp

@@
 First, the frequency response of the filters in the first layer which operates on the raw waveform is plotted.
-As the order of learned filters in a convolutional layer is arbitrary, we sort the filters by the peak value of the frequency response.
+As the order of learned filters in a convolutional layer is arbitrary, we sort the filters by the peak value of the frequency response and by the upper and lower \SI{3}{\decibel} cutoff frequencies as second and third criteria, respectively.
 The plots are depicted in \refFig{fig:first_layer}\,(a)-(d), which share the same color range.
@@
 The distribution resembles the Gammatone filterbank and with higher center frequencies, the pass band is also wider.
-However, the stopband attenuation is clearly weaker.
+However, the stopband attenuation is clearly weaker and there is also a number of filters with no clear passband and a rather uncharacteristic frequency response.
+This is particularly visible in \refFig{fig:first_layer}\,(e).
 
@@
 Again, bandpass filters are learned, however, they have a wider passband, their center frequencies are distributed rather linearly and the stopband attenuation is weak.
+In addition, about 14\% of all filters have their peak at \SI{0}{\kilo\hertz} and another 10\% at \SI{8}{\kilo\hertz} with varying bandwidths.
+Moreover, a number of filters exists with multiple passbands.
 Interestingly, the frequency response of the pre-trained \wvtwo \fe in \refFig{fig:first_layer}\,(d) is very similar to \refFig{fig:first_layer}\,(c).
 This indicates that the properties learned in unsupervised pre-training and supervised \gls{CTC} training are similar.
-% for SCF, first 9 filters have peak at 0Hz, for wav2vec 75 filters
 

### 2023-03-09 main.tex
update figues, add text for full fe
fix arxiv refs
add text for masking exps

@@
 \subsection{Frequency Response of Learned Filters}
+\label{sec:freq_resp}
 Finally, we analyze the learned filters of the neural \fes.
@@
 
+\refFig{fig:first_layer}\,(e) and (f) show the frequency response of the complete \fe.
+For \wvtwo, the distribution of center frequencies is not as linear anymore and as suggested in \cite{choi2022w2v2fe}.
+Moreover, the ratio of peak to mean values is much lower than for the \gls{SC} features.
+
+\subsection{Filter Masking}
+The observed uncharacteristic wideband filters from \refSec{sec:freq_resp} raise the question whether their output is particularly helpful for the model in addition to the bandpass filters or if they are just not trained properly and ignored by the system.
+To answer this question, we masked a portion of the 150 filters in the \gls{SC} features' first layer.
+The filters are sorted by the peak-to-average ratio of their frequency response and the $N$ filters with the lowest ratio are masked out and replaced by zeros.
+\TODO{Show results}.
+We can observe that masking the first 20 filters has no effect on the \gls{WER} and only after masking a 50, so a third of all filters, the \gls{WER} degrades significantly.
+This demonstrates that the wideband filters do not contain valuable information for the \gls{ASR} system and are rather the result of a suboptimal training process.
+Future work should address this observation to allow a better exploitation of the \fe's potential.
+
 \section{Conclusions}
@@
 
-{
-\color{red}
-\section{To Dos/Questions}
-\begin{itemize}
-  \item check for arxiv citations
-\end{itemize}
-}
-
 \bibliographystyle{IEEEtran}

### 2023-03-09 main.tex
add figure for masking

@@
 \subsection{Filter Masking}
+\input{figures/masked_filters}
 The observed uncharacteristic wideband filters from \refSec{sec:freq_resp} raise the question whether their output is particularly helpful for the model in addition to the bandpass filters or if they are just not trained properly and ignored by the system.
@@
 The filters are sorted by the peak-to-average ratio of their frequency response and the $N$ filters with the lowest ratio are masked out and replaced by zeros.
-\TODO{Show results}.
-We can observe that masking the first 20 filters has no effect on the \gls{WER} and only after masking a 50, so a third of all filters, the \gls{WER} degrades significantly.
+In \refFig{fig:masked_filters}, we can observe that masking the first 20 filters has no effect on the \gls{WER} and only after masking a 50, so a third of all filters, the \gls{WER} degrades significantly.
 This demonstrates that the wideband filters do not contain valuable information for the \gls{ASR} system and are rather the result of a suboptimal training process.

### 2023-03-09 main.tex
revise abstract

@@
 \begin{abstract}
-Raw waveform modeling is an interesting research topic for \gls{ASR}.
-\wvtwo has seen large success recently and makes use of a convolutional \fe which operates directly on the speech waveform.
-Typically, this \fe is seen as part of the overall architecture, but this work investigates this component specifically.
-It is used as a replacement for classical feature extraction methods in a \gls{CTC} setup and compared to an alternative neural \fe approach.
-Finally, we analyze the learned representations of both neural \fes.
-\TODO{Revise abstract and include results.}
+\Gls{ASR} systems typically use handcrafted feature extraction pipelines.
+To avoid their inherent information loss and to achieve a more consistent modeling from speech to transcribed text, neural raw waveform \fes are an appealing approach.
+Also the \wvtwo model, which has recently gained large popularity, uses a convolutional \fe which operates directly on the speech waveform.
+However, it is not yet studied extensively in the literature.
+In this work, we study its capability to replace the standard feature extraction methods in a \gls{CTC} \gls{ASR} model and compare it to an alternative neural \fe.
+We show that both are competitive with traditional \fes on the LibriSpeech benchmark and analyze the effect of the individual components.
+Furthermore, we analyze the learned filters and show that the most important information for the \gls{ASR} system is obtained by a set of bandpass filters.
 \end{abstract}

### 2023-03-09 main.tex
add contributions text

@@
 
-\draft{
-Highlight contributions of this paper.
-}
+In this work, we show new insights into the \wvtwo \fe.
+Unlike previous work, we demonstrate that it can be used as a replacement for traditional feature extraction methods for a \gls{CTC} model similar to \gls{SC} features.
+The results show that both are competitive on the LibriSpeech benchmark.
+We analyze the learned filters and show the differences between the characteristics of both neural \fes.
+Finally, we demonstrate that the learned bandpass filters of the \gls{SC} \fe encode the valuable information for the \gls{ASR} system while the remaining wideband filter information is largely ignored.
 
@@
 \subsubsection{Supervised Convolutional Features}
-As an alternative comparable neural feature extraction method, we use \gls{SC} features \cite{tuske2018:waveform, vieting2021waveform}.
+As an alternative comparable neural feature extraction method, we use the \gls{SC} features from \cite{tuske2018:waveform, vieting2021waveform}.
 Similarly, a convolutional filterbank with learnable parameters is applied to the waveform.

### 2023-03-09 main.tex
remove ctc todo, add no more information
small fixes/updates

@@
 It follows the setup described for the \gls{CTC} model in \cite{zhou2022efficient}.
-\TODO{Add more description of the CTC model.}
 All models are trained in a purely supervised fashion using the \gls{CTC} objective unless explicitly stated otherwise.
@@
 It can be observed that these configurations perform similarly to the 6 layer 512-dim baseline while reducing the training time by about 21\%, 19\% and 14\%, respectively.
+Note that the \gls{SC} features perform similar to the variant with dimension 64-512 while again around 10\% shorter training time.
 

### 2023-03-09 tables/features_pretraining.tex
small fixes/updates

--- 2023-03-07 tables/features_pretraining.tex
@@
                  \multicolumn{2}{|c|}{} &      \multicolumn{2}{c|}{{dev}} & \multicolumn{2}{c|}{{test}} \\\hline
-                              Trainable & Pretrained &                         {clean} & other &                     {clean} & other \\\hline\hline
-                                    yes &         no &                             2.9 &   6.8 &                             &   7.5 \\\cline{2-6}
-                                        &        yes &                             2.9 &   6.9 &                             &   7.5 \\\cline{1-1}\cline{3-6}
-                                     no &            &                             3.0 &   7.1 &                             &   7.8 \\
+                              Trainable & Pre-trained &                         {clean} & other &                     {clean} & other \\\hline\hline
+                                    yes &          no &                             2.9 &   6.8 &                         3.3 &   7.5 \\\cline{2-6}
+                                        &         yes &                             2.9 &   6.9 &                         3.3 &   7.5 \\\cline{1-1}\cline{3-6}
+                                     no &             &                             3.0 &   7.1 &                         3.3 &   7.8 \\
 \hline

### 2023-03-09 main.tex
add conclusion

@@
 \section{Conclusions}
+In this work, we utilize the \wvtwo \fe as a replacement for traditional feature extraction methods in a \gls{CTC} \gls{ASR} model on LibriSpeech.
+It shows competitive performance and outperforms an alternative neural \fe, namely the \gls{SC} features, slightly.
+We demonstrate that its width is more responsible for its capabilities that its depth and show that pre-training the \fe only is not beneficial.
+Finally, we analyze the learned filters and highlight the differences between both neural \fes.
+In particular, it is shown that the \gls{SC} \fe's learned bandpass filters encode the valuable information for the \gls{ASR} system while the remaining wideband filter information is largely ignored.
 

### 2023-03-09 main.tex
proofreading

@@
 We can observe that all \fes are within a close range.
-The \gls{SC} features slightly outperform the Gammatone baseline, while log Mel features are again marginally better and the \wvtwo \fe yields the best results.
+The \gls{SC} features slightly outperform the Gammatone baseline, while log Mel features and the \wvtwo \fe are again better with advantages on \devother and \testother, respectively.
 This demonstrates that learnable neural \fes are competitive with hand-designed methods in a purely supervised setup with 960h of training data.
 However, a clear advantage of these techniques cannot be deduced from the experimental results.
+
 The number of parameters for Gammatone and log Mel features refers to the \gls{FIR} and Mel filterbanks, respectively, that we use in our implementation even if they are not trainable.
@@
 The kernel sizes and strides were chosen as (10, 3, 3, 3, 3, 2), (5, 2, 2, 2, 2, 2) for 6 layers, (10, 6, 3, 3, 3), (5, 4, 2, 2, 2) for 5 layers, (10, 6, 6, 3), (5, 4, 4, 2) for 4 layers, (20, 6, 6), (10, 4, 4) for 3 layers and (32, 20), (16, 10) for 2 layers.
-The results show that a larger inner dimension leads to a better performance
+The results show that decreasing the inner dimension deteriorates the performance, however, a larger dimension does not improve over the 512-dim baseline.
 In contrast, no major effect can be observed regarding the number of layers.
@@
 It can be observed that these configurations perform similarly to the 6 layer 512-dim baseline while reducing the training time by about 21\%, 19\% and 14\%, respectively.
-Note that the \gls{SC} features perform similar to the variant with dimension 64-512 while again around 10\% shorter training time.
+Note that the \gls{SC} features perform similar to the variant with dimension 64-512 with again around 10\% shorter training time.
 
@@
 They suggest that smaller window sizes than in \cite{tuske2018:waveform} perform better in the context of a \conformer \gls{CTC} model.
-In particular, it does not seem to be necessary to cover one full pitch period which is around \ms{16} for low voices.
+In particular, it does not seem to be necessary to cover one full pitch period which can be around \ms{16} for low voices.
 Instead, the best result is achieved with a window size of \ms{10} and even smaller windows still achieve comparable performance.
@@
 \cite{facebook2020wav2vec2} even reports relative \gls{WER} improvements of about 10\% on the \textit{other} subsets for their experiments on LibriSpeech when pre-training on the same data as used in supervised training compared to pure supervised training from scratch.
-Here, we study this effect for the \fe only by using the parameters of an \fe pre-trained on LibriSpeech\footnote{\raggedright\url{https://github.com/facebookresearch/fairseq/blob/main/examples/wav2vec/README.md}}, so no external data is added.
+Here, we study this effect only for the \fe by using the parameters of an \fe pre-trained on LibriSpeech\footnote{\raggedright\url{https://github.com/facebookresearch/fairseq/blob/main/examples/wav2vec/README.md}}, so no external data is added.
 During the supervised training, the \fe weights can be further trained or kept frozen.
@@
 \label{sec:freq_resp}
-Finally, we analyze the learned filters of the neural \fes.
-First, the frequency response of the filters in the first layer which operates on the raw waveform is plotted.
+To analyze the learned filters of the neural \fes, we first plot the frequency response of the filters in the first layer which operates on the raw waveform.
 As the order of learned filters in a convolutional layer is arbitrary, we sort the filters by the peak value of the frequency response and by the upper and lower \SI{3}{\decibel} cutoff frequencies as second and third criteria, respectively.
@@
 In line with \cite{tuske2018:waveform}, bandpass filters with a non-linear distribution of center frequencies are learned.
-This confirms, that these filters are also preferred by more advanced neural back-ends like self-attention-based like \conformer which are superior in the exploitation of context information compared to simpler neural back-ends e.g.
+This confirms, that these filters are also preferred by more advanced neural back-ends like the self-attention-based \conformer which are superior in the exploitation of context information compared to simpler neural back-ends e.g.
 in \cite{tuske2014raw}.
-The distribution resembles the Gammatone filterbank and with higher center frequencies, the pass band is also wider.
+The distribution resembles the Gammatone filterbank and with higher center frequencies, the passband is also wider.
 However, the stopband attenuation is clearly weaker and there is also a number of filters with no clear passband and a rather uncharacteristic frequency response.
-This is particularly visible in \refFig{fig:first_layer}\,(e).
 
@@
 To answer this question, we masked a portion of the 150 filters in the \gls{SC} features' first layer.
-The filters are sorted by the peak-to-average ratio of their frequency response and the $N$ filters with the lowest ratio are masked out and replaced by zeros.
-In \refFig{fig:masked_filters}, we can observe that masking the first 20 filters has no effect on the \gls{WER} and only after masking a 50, so a third of all filters, the \gls{WER} degrades significantly.
+The filters are sorted by the peak-to-average ratio of their frequency response and the $N$ filters with the lowest ratio are masked out during recognition.
+In \refFig{fig:masked_filters}, we can observe that masking the first 20 filters has no effect on the \gls{WER} and only after masking 50, so a third of all filters, the \gls{WER} degrades significantly.
 This demonstrates that the wideband filters do not contain valuable information for the \gls{ASR} system and are rather the result of a suboptimal training process.

### 2023-03-09 main.tex
cut sections on w2v proj, scf size and scf window

@@
 
-\subsection{Model Sizes}
+\subsection{Dissecting the wav2vec 2.0 Feature Extractor}
+\label{sec:w2v_components}
 It is striking that the \wvtwo \fe uses two orders of magnitude more parameters than the \gls{SC} architecture.
-We therefore study the contribution of its components in \refSec{sec:w2v_components} and increase the number of parameters in the \gls{SC} model in \refSec{sec:scf_size}.
-Additionally, the size of the first window operating on the waveform is analyzed in \refSec{sec:scf_first_window}.
-
-\subsubsection{Dissecting the wav2vec 2.0 Feature Extractor}
-\label{sec:w2v_components}
-To understand where \wvtwo uses the parameters and how much they contribute to the \gls{FE}'s performance, we run it with different configurations.
+To understand where \wvtwo uses the them and how much they contribute to the \gls{FE}'s performance, we run it with different configurations.
 The results for different widths and depths are shown in \refTab{table:features_w2v_size}.
@@
 We also tried a variant with only one layer using kernel size 320 and stride 160 which did not converge.
+\input{tables/features_w2v_size}
+Removing the final projection with about 394k parameters has no significant impact.
+%The final projection, which comprises about 394k parameters, additionally improves the performance slightly on \devother while degrading minimally on the \textit{test} sets as shown in \refTab{table:features_w2v_proj}.
+%\input{tables/features_w2v_proj}
 
@@
 Note that the \gls{SC} features perform similar to the variant with dimension 64-512 with again around 10\% shorter training time.
-
-The final projection, which comprises about 394k parameters, additionally improves the performance slightly on \devother while degrading minimally on the \textit{test} sets as shown in \refTab{table:features_w2v_proj}.
-\input{tables/features_w2v_size}
-\input{tables/features_w2v_proj}
 % has to be here to be shown on page 4
@@
 
-\subsubsection{Larger Supervised Convolutional Architecture}
-\label{sec:scf_size}
-As a larger inner dimension clearly showed better performance for the \wvtwo \gls{FE}, we increase the inner dimensions of the \gls{SC} features as well.
-However, as shown in \refTab{table:features_scf_size}, neither increasing the number of filters operating on the waveform nor the number of kernels for the multi-resolutional temporal integration helps to improve the performance.
-Note that the number of parameters of the full model is even larger than for the \wvtwo model in \refTab{table:features_general}, however, a large portion of those parameters is consumed by the linear projection of the features after the \gls{VGG} block.
-\input{tables/features_scf_size}
+%\subsubsection{Larger Supervised Convolutional Architecture}
+%\label{sec:scf_size}
+%As a larger inner dimension clearly showed better performance for the \wvtwo \gls{FE}, we increase the inner dimensions of the \gls{SC} features as well.
+%However, as shown in \refTab{table:features_scf_size}, neither increasing the number of filters operating on the waveform nor the number of kernels for the multi-resolutional temporal integration helps to improve the performance.
+%Note that the number of parameters of the full model is even larger than for the \wvtwo model in \refTab{table:features_general}, however, a large portion of those parameters is consumed by the linear projection of the features after the \gls{VGG} block.
+%\input{tables/features_scf_size}
 
-\subsubsection{First Layer Window Size}
-\label{sec:scf_first_window}
-Furthermore, the effect of the first layer's window size is studied and the results are shown in \refTab{table:features_window_size}.
-They suggest that smaller window sizes than in \cite{tuske2018:waveform} perform better in the context of a \conformer \gls{CTC} model.
-In particular, it does not seem to be necessary to cover one full pitch period which can be around \ms{16} for low voices.
-Instead, the best result is achieved with a window size of \ms{10} and even smaller windows still achieve comparable performance.
-Note that all experiments in \refSec{sec:scf_size} use a window size of \ms{16}.
-\input{tables/features_window_size}
+%\subsection{First Layer Window Size}
+%\label{sec:scf_first_window}
+%Furthermore, the effect of the first layer's window size is studied and the results are shown in \refTab{table:features_window_size}.
+%They suggest that smaller window sizes than in \cite{tuske2018:waveform} perform better in the context of a \conformer \gls{CTC} model.
+%In particular, it does not seem to be necessary to cover one full pitch period which can be around \ms{16} for low voices.
+%Instead, the best result is achieved with a window size of \ms{10} and even smaller windows still achieve comparable performance.
+%Note that all experiments in \refSec{sec:scf_size} use a window size of \ms{16}.
+%\input{tables/features_window_size}
 

### 2023-03-09 tables/create_tables.sh
squeeze space in table creation

@@ -14,3 +14,5 @@ python ${EXCEL2LATEX} --file_in features.xlsx --vertical_merge --sheet scf_size 
 python ${EXCEL2LATEX} --file_in features.xlsx --all_lines --vertical_merge --sheet w2v_size --wer_cols D E F G
 python ${EXCEL2LATEX} --file_in features.xlsx --all_lines --vertical_merge --sheet w2v_proj --wer_cols C D E F
 
+sed -i 's/\\end{tabular}/\\end{tabular}\n\\vspace{-0.05cm}/g' features_w2v_size.tex
+

### 2023-03-09 main.tex
squeeze space
make blind for review
nitpicks

@@
 %  'camera ready' version to submit for publication.
-\interspeechcameraready
+% \interspeechcameraready
 
@@
 \subsection{Acoustic Model}
-The \gls{AM} mainly consists of \gls{VGG} \addref blocks for downsampling and 12 \conformer blocks.
+The \gls{AM} mainly consists of 3 VGG blocks for downsampling and 12 \conformer blocks.
 Its configuration mostly follows \cite{zeineldeen2022conformer} and \cite{zhou2022efficient}.
-The \gls{VGG} downsampling uses three 3x3-convolutional layers with 32, 64 and 64 channels, respectively.
+The VGG downsampling uses three 3x3-convolutional layers with 32, 64 and 64 channels, respectively.
 In total, a subsampling factor of 4 in the time dimension is achieved by the strided convolutions so that the \conformer operates on frames with a shift of \SI{40}{\milli\second}.
@@
 \input{tables/features_general}
+\vspace{-0.3cm}
 
@@
 We also tried a variant with only one layer using kernel size 320 and stride 160 which did not converge.
-\input{tables/features_w2v_size}
 Removing the final projection with about 394k parameters has no significant impact.
 %The final projection, which comprises about 394k parameters, additionally improves the performance slightly on \devother while degrading minimally on the \textit{test} sets as shown in \refTab{table:features_w2v_proj}.
+\input{tables/features_w2v_size}
 %\input{tables/features_w2v_proj}
@@
 To analyze the learned filters of the neural \fes, we first plot the frequency response of the filters in the first layer which operates on the raw waveform.
-As the order of learned filters in a convolutional layer is arbitrary, we sort the filters by the peak value of the frequency response and by the upper and lower \SI{3}{\decibel} cutoff frequencies as second and third criteria, respectively.
+As the order of learned filters in a convolutional layer is arbitrary, we sort the filters by the peak value of the frequency response and by the upper and lower \SI{3}{\decibel} cutoff frequencies as second and third criteria.
 The plots are depicted in \refFig{fig:first_layer}\,(a)-(d), which share the same color range.

### 2023-03-09 main.tex
add rasr/returnn, squeeze

@@
 Each model is trained on a single \gls{GPU}.
+The experiments are carried out using 
+\ifinterspeechfinal
+     RETURNN and RASR
+\else
+     [\textit{toolkits blind for review}]
+\fi
+and configs will be made available at
+\ifinterspeechfinal
+     https://github.com/rwth-i6/.
+\else
+     [\textit{blind for review}].
+\fi
 
@@
 \input{tables/features_general}
-\vspace{-0.3cm}
 
@@
 In fact, using a frozen pre-trained \fe shows clearly worse performance than training the \fe purely supervised from scratch.
-We hypothesize that the pre-training mainly benefits the \transformer encoder and not the \fe while the mismatch between the \transformer encoder in pre-training and \conformer encoder in our setup could also pose a challenge for the case that the \fe is frozen.
+We hypothesize that the pre-training mainly benefits the \transformer encoder and not the \fe while the mismatch between the \transformer encoder in pre-training and \conformer encoder in our setup could also pose a challenge for the frozen \fe.
 \input{tables/features_pretraining}
 
+\vspace{-0.5cm}
 \subsection{Frequency Response of Learned Filters}
@@
 In addition, about 14\% of all filters have their peak at \SI{0}{\kilo\hertz} and another 10\% at \SI{8}{\kilo\hertz} with varying bandwidths.
-Moreover, a number of filters exists with multiple passbands.
+A number of filters exists with multiple passbands.
 Interestingly, the frequency response of the pre-trained \wvtwo \fe in \refFig{fig:first_layer}\,(d) is very similar to \refFig{fig:first_layer}\,(c).
@@
 \section{Acknowledgements}
-
 \ifinterspeechfinal

### 2023-03-15 main.tex
fix typos and grammar with grammarly

@@
 \Gls{ASR} systems typically use handcrafted feature extraction pipelines.
-To avoid their inherent information loss and to achieve a more consistent modeling from speech to transcribed text, neural raw waveform \fes are an appealing approach.
+To avoid their inherent information loss and to achieve more consistent modeling from speech to transcribed text, neural raw waveform \fes are an appealing approach.
 Also the \wvtwo model, which has recently gained large popularity, uses a convolutional \fe which operates directly on the speech waveform.
@@
 These techniques are motivated by insights into the properties of the human auditory system.
-However, their fixed design inevitably leads to information loss during the feature extraction.
+However, their fixed design inevitably leads to information loss during feature extraction.
 
@@
 Since their parameters are learned during training, these \fes can adapt to the need of \glspl{AM} and the loss of particularly helpful information may be avoided.
-A number of works have been proposed in this direction, often using convolutional \glspl{NN} \cite{palaz2015convolutional, golik15:cnn, tuske2018:waveform} but also applying parametrized functions \cite{ravanelli2018sincnet} and other architectures \cite{sainath2015cldnn}.
+Several works have been proposed in this direction, often using convolutional \glspl{NN} \cite{palaz2015convolutional, golik15:cnn, tuske2018:waveform} but also applying parametrized functions \cite{ravanelli2018sincnet} and other architectures \cite{sainath2015cldnn}.
 
@@
 In particular, different neural \fes are compared to traditional feature extraction methods.
-We choose to investigate the supervised convolutional \fe from \cite{tuske2018:waveform}, that was also studied in \cite{vieting2021waveform}.
+We choose to investigate the supervised convolutional \fe from \cite{tuske2018:waveform}, which was also studied in \cite{vieting2021waveform}.
 Furthermore, the feature extractor of \wvtwo \cite{facebook2020wav2vec2} is used as a learnable front-end for supervised \gls{ASR} training.
@@
 %
-Also the \fe itself has been adjusted.
+Also, the \fe itself has been adjusted.
 %\cite{facebook2019wav2vec} uses 5 layers with kernel sizes (10, 8, 4, 4, 4) and strides which are always exactly half the kernel size.
@@
 
-While most works analyzing the \wvtwo model focus on the \transformer part, there are papers which study the \fe as well.
+While most works analyzing the \wvtwo model focus on the \transformer part, some papers study the \fe as well.
 The authors of \cite{choi2022w2v2fe} feed synthesized sine waves to a pre-trained model and observe that enough temporal detail is obtained and information about fundamental frequencies is represented in the model's output.
@@
 \subsubsection{Standard Feature Extraction}
-As a reference, we use standard, hand-designed \fes -- namely log Mel filterbank features and Gammatone features.
-To compute the log Mel filterbank features, first the \gls{STFT} of the waveform is computed with a window size of \ms{25} and a shift of \ms{10}.
-We keep the square of the magnitude and perform Mel warping to obtain a 80-dim vector.
+As a reference, we use standard, hand-designed \fes -- namely, log Mel filterbank features and Gammatone features.
+To compute the log Mel filterbank features, first, the \gls{STFT} of the waveform is computed with a window size of \ms{25} and a shift of \ms{10}.
+We keep the square of the magnitude and perform Mel warping to obtain an 80-dim vector.
 Finally, $\log_{10}$ and normalization are applied.
@@
 Gammatone features apply a Gammatone filterbank to the pre-emphasized speech signal, perform temporal integration over each channel using a low pass filter, i.e., the Hanning window with a size of \ms{25} and a shift of \ms{10}, compress using the $10^{th}$ root and finally compute the \gls{DCT} of the values \cite{schlueter:icassp07}.
-Further, the the resulting 50-dim features are normalized.
+Further, the resulting 50-dim features are normalized.
 
@@
 They are configured with kernel sizes (10, 3, 3, 3, 3, 2, 2), strides (5, 2, 2, 2, 2, 2, 2), and a GELU activation function.
-The first layer applies group normalization and finally, layer normalization and a linear projection is added.
+The first layer applies group normalization and finally, layer normalization and a linear projection are added.
 The total receptive field of the \wvtwo \fe is \ms{25} and the shift between consecutive frames is \ms{20}.
-Since the other \fes operate at a shift of \ms{10} and the \wvtwo \fe is built in a modular way, we remove the last layer with a stride of 2 in order to achieve features at the same frame rate.
+Since the other \fes operate at a shift of \ms{10} and the \wvtwo \fe is built in a modular way, we remove the last layer with a stride of 2 to achieve features at the same frame rate.
 This reduces the total receptive field to \ms{15}.
@@
 As in the case of Gammatones, a temporal integration is performed over each channel.
-However, in this case multiple filters are used for temporal integration allowing for multi-resolutional processing.
+However, in this case, multiple filters are used for temporal integration allowing for multi-resolutional processing.
 Additionally, these filters are learned during training.
@@
 For the experiments using a pre-trained \wvtwo \fe, the same 960h were used in pre-training.
-The \gls{LM} was trained on the default LibriSpeech \gls{LM} training data which includes the transcriptions of the training set along with the additional 800M-word text only data.
+The \gls{LM} was trained on the default LibriSpeech \gls{LM} training data which includes the transcriptions of the training set along with the additional 800M-word text-only data.
 The evaluation is done on the standard \devclean, \devother, \testclean and \testother sets.
@@
 We generally follow the setup in \cite{zhou2022efficient}.
-The batch size is 640k samples, which corresponds to \SI{40}{\second} of speech waveform and the gradients are accumulated over 3 steps.
+The batch size is 640k samples, which corresponds to \SI{40}{\second} of the speech waveform and the gradients are accumulated over 3 steps.
 The NAdam optimizer uses a \gls{OCLR} schedule with a peak learning rate of $3\cdot 10^{-4}$ and we train the model for 20 epochs.
@@
 It is striking that the \wvtwo \fe uses two orders of magnitude more parameters than the \gls{SC} architecture.
-To understand where \wvtwo uses the them and how much they contribute to the \gls{FE}'s performance, we run it with different configurations.
+To understand where \wvtwo uses them and how much they contribute to the \gls{FE}'s performance, we run it with different configurations.
 The results for different widths and depths are shown in \refTab{table:features_w2v_size}.
@@
 Additionally, we add point-wise convolutional layers after all but the first layer using the same inner dimension as the preceding layer for the last row of the table.
-It can be observed that these configurations perform similarly to the 6 layer 512-dim baseline while reducing the training time by about 21\%, 19\% and 14\%, respectively.
-Note that the \gls{SC} features perform similar to the variant with dimension 64-512 with again around 10\% shorter training time.
+It can be observed that these configurations perform similarly to the 6-layer 512-dim baseline while reducing the training time by about 21\%, 19\% and 14\%, respectively.
+Note that the \gls{SC} features perform similarly to the variant with dimension 64-512 with again around 10\% shorter training time.
 % has to be here to be shown on page 4
@@
 The main appeal of the \wvtwo framework is the ability to pre-train the model on unlabeled audio data.
-\cite{facebook2020wav2vec2} even reports relative \gls{WER} improvements of about 10\% on the \textit{other} subsets for their experiments on LibriSpeech when pre-training on the same data as used in supervised training compared to pure supervised training from scratch.
+\cite{facebook2020wav2vec2} even reports relative \gls{WER} improvements of about 10\% on the \textit{other} subsets for their experiments on LibriSpeech when pre-training on the same data as used in supervised training compared to purely supervised training from scratch.
 Here, we study this effect only for the \fe by using the parameters of an \fe pre-trained on LibriSpeech\footnote{\raggedright\url{https://github.com/facebookresearch/fairseq/blob/main/examples/wav2vec/README.md}}, so no external data is added.
@@
 
-The well known distribution of the Gammatone filters is given in \refFig{fig:first_layer}\,(a) as a reference.
+The well-known distribution of the Gammatone filters is given in \refFig{fig:first_layer}\,(a) as a reference.
 \refFig{fig:first_layer}\,(b) shows the frequency response of the learned \gls{SC} filters.
@@
 The distribution resembles the Gammatone filterbank and with higher center frequencies, the passband is also wider.
-However, the stopband attenuation is clearly weaker and there is also a number of filters with no clear passband and a rather uncharacteristic frequency response.
+However, the stopband attenuation is clearly weaker and there are also several filters with no clear passband and a rather uncharacteristic frequency response.
 
@@
 In addition, about 14\% of all filters have their peak at \SI{0}{\kilo\hertz} and another 10\% at \SI{8}{\kilo\hertz} with varying bandwidths.
-A number of filters exists with multiple passbands.
+Several filters exist with multiple passbands.
 Interestingly, the frequency response of the pre-trained \wvtwo \fe in \refFig{fig:first_layer}\,(d) is very similar to \refFig{fig:first_layer}\,(c).
@@
 \input{figures/masked_filters}
-The observed uncharacteristic wideband filters from \refSec{sec:freq_resp} raise the question whether their output is particularly helpful for the model in addition to the bandpass filters or if they are just not trained properly and ignored by the system.
+The observed uncharacteristic wideband filters from \refSec{sec:freq_resp} raise the question of whether their output is particularly helpful for the model in addition to the bandpass filters or if they are just not trained properly and ignored by the system.
 To answer this question, we masked a portion of the 150 filters in the \gls{SC} features' first layer.
@@
 This demonstrates that the wideband filters do not contain valuable information for the \gls{ASR} system and are rather the result of a suboptimal training process.
-Future work should address this observation to allow a better exploitation of the \fe's potential.
+Future work should address this observation to allow better exploitation of the \fe's potential.
 
@@
 \else
-     The acknowledgements would likely reveal the author identity and are therefore hidden for the double-blind review process.
+     The acknowledgements would likely reveal the authors' identity and are therefore hidden for the double-blind review process.
 \fi

### 2023-05-03 rebuttal.tex
add rebuttal

\documentclass[11pt]{extarticle}
\usepackage[backend=biber,style=alphabetic,]{biblatex}
\usepackage[a4paper, margin=1in]{geometry}
\usepackage[utf8]{inputenc}
\usepackage{layout}
\usepackage{siunitx}
\usepackage{xcolor}

\title{Rebuttal on the Reviews for the Interspeech 2023 Submission \\ "Put Title Here"}
% \author{}
\date{}

\pagenumbering{gobble}

\begin{document}
\maketitle\vspace{-10mm}
First of all, we thank all the reviewers for their time and valuable feedback! Their responses are appreciated and we identified the following points on which we would like to comment:

\paragraph{Motivation}
\begin{quote}\textit{
Light-weight audio frontends represent an interesting track to replace classic mel spectrograms, but W2V2 FE is very heavy compared to the considered competitors here, and it does not seem to yield interesting results.
\[...\]
\[The\] feature encoder in wav2vec2.0 is generally not updated during the supervised fine-tuning \[...\] and are only trained then with the unsupervised loss.
\[...\]
I think from this study, that the interesting question would be whether this front-end is really needed for W2V2 \[...\].
--- Reviewer~1
}\end{quote}
Light-weight frontends are an interesting approach, however, recent trends in ASR models tend towards larger models which clearly outperform smaller models.
Thus, a natural question is whether larger audio frondends are helpful in a similar way.

It is correct that in wav2vec 2.0, the feature extractor is only trained with the unsupervised loss.
We study the effect both in Table~3 and Figure~1 and show (for the first time to the best of our knowledge), that similar results and representations are achieved for both an unsupervised and a supervised loss.

We agree that this work raises the question whether the learnable front-end is actually needed in wav2vec 2.0 and related approaches.
However, this does not contradict the motivation of this work but instead makes it a valuable contribution to the ongoing research on wav2vec-like models.

\paragraph{Baseline Performance}
\begin{quote}\textit{
The main drawback here is that the results do look a little bit too poor compared to the literature, and to the papers originally implementing some of the baseline techniques.
For instance, the conformer paper which also builds upon Mel Filterbanks reaches a performance of 2.0/4.3 on librispeech test splits using the official language model, in the medium version containing 30M parameters.
--- Reviewer~1
}\end{quote}
As the baseline, we use the CTC model from \cite{zhou2022efficient} with some modifications mentioned in the paper which lead to an improvement compared to the original paper (from 7.6\% to 7.1\% on LibriSpeech dev-other).
This CTC model is part of the pipeline which finally yields a neural transducer model with a competitive performance of 2.1\%/4.1\% on LibriSpeech test-clean and test-other.
While we acknowledge that the CTC results are inferior to this final performance, we expect the relative trends to transfer.
The 2.0\%/4.3\% in \cite{google2020conformer} mentioned by Reviewer~1 are comparable to the results in \cite{zhou2022efficient} but are not directly comparable to our setup because a neural LSTM LM is used (not the official 4-gram LM as claimed by the reviewer, which we use here and which is much weaker) and \cite{google2020conformer} uses a transducer while we use a CTC without label feedback.
We expect all these points to be orthogonal to the investigation of the feature extractor done here.

\paragraph{Reproducibility / Open Codebase}
\begin{quote}\textit{
\[...\] the absence of any codebase does not allow us to inspect the reasons behind the difference in terms of results.
--- Reviewer~1
}\end{quote}
We generally make our recipes available and this also holds for the work proposed here, however, the reference would only be added in the camera ready version as it would have revealed the author identity during the review process.

\paragraph{Clarification}
\begin{quote}\textit{
Not clear what LM have they used here.
--- Reviewer~4
}\end{quote}
This is explicitly stated in Section 3.3: We use the official LibriSpeech 4-gram LM.

\bibliographystyle{IEEEtran}
\bibliography{mybib}

\end{document}

### 2023-05-03 rebuttal.tex
updates to rebuttal

@@
 \documentclass[11pt]{extarticle}
-\usepackage[backend=biber,style=alphabetic,]{biblatex}
 \usepackage[a4paper, margin=1in]{geometry}
@@
 
-\title{Rebuttal on the Reviews for the Interspeech 2023 Submission \\ "Put Title Here"}
+\title{Rebuttal on the Reviews for the Interspeech 2023 Submission \\ "Comparative Analysis of the wav2vec 2.0 Feature Extractor"}
 % \author{}
@@
 
+\setlength{\voffset}{-0.75in}
+\setlength{\hoffset}{-25pt}
+\setlength{\textheight}{800pt}
+\setlength{\textwidth}{500pt}
 \pagenumbering{gobble}
@@
 Light-weight audio frontends represent an interesting track to replace classic mel spectrograms, but W2V2 FE is very heavy compared to the considered competitors here, and it does not seem to yield interesting results.
-\[...\]
-\[The\] feature encoder in wav2vec2.0 is generally not updated during the supervised fine-tuning \[...\] and are only trained then with the unsupervised loss.
-\[...\]
-I think from this study, that the interesting question would be whether this front-end is really needed for W2V2 \[...\].
+[...]
+[The] feature encoder in wav2vec2.0 is generally not updated during the supervised fine-tuning [...] and are only trained then with the unsupervised loss.
+[...]
+I think from this study, that the interesting question would be whether this front-end is really needed for W2V2 [...].
 --- Reviewer~1
 }\end{quote}
-Light-weight frontends are an interesting approach, however, recent trends in ASR models tend towards larger models which clearly outperform smaller models.
+Recent trends in ASR models tend towards ever larger models.
 Thus, a natural question is whether larger audio frondends are helpful in a similar way.
-
 It is correct that in wav2vec 2.0, the feature extractor is only trained with the unsupervised loss.
 We study the effect both in Table~3 and Figure~1 and show (for the first time to the best of our knowledge), that similar results and representations are achieved for both an unsupervised and a supervised loss.
-
 We agree that this work raises the question whether the learnable front-end is actually needed in wav2vec 2.0 and related approaches.
@@
 The main drawback here is that the results do look a little bit too poor compared to the literature, and to the papers originally implementing some of the baseline techniques.
-For instance, the conformer paper which also builds upon Mel Filterbanks reaches a performance of 2.0/4.3 on librispeech test splits using the official language model, in the medium version containing 30M parameters.
 --- Reviewer~1
@@
 While we acknowledge that the CTC results are inferior to this final performance, we expect the relative trends to transfer.
-The 2.0\%/4.3\% in \cite{google2020conformer} mentioned by Reviewer~1 are comparable to the results in \cite{zhou2022efficient} but are not directly comparable to our setup because a neural LSTM LM is used (not the official 4-gram LM as claimed by the reviewer, which we use here and which is much weaker) and \cite{google2020conformer} uses a transducer while we use a CTC without label feedback.
+The 2.0\%/4.3\% in the conformer paper mentioned by Reviewer~1 are comparable to the results in \cite{zhou2022efficient} but are not directly comparable to our setup because a neural LSTM LM is used and the conformer paper uses a transducer while we use a CTC without label feedback.
 We expect all these points to be orthogonal to the investigation of the feature extractor done here.
@@
 \begin{quote}\textit{
-\[...\] the absence of any codebase does not allow us to inspect the reasons behind the difference in terms of results.
+[The] absence of any codebase does not allow us to inspect the reasons behind the difference in terms of results.
 --- Reviewer~1
